### 2021-03-04 article.tex
first draft

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{The impacts of the beginning of the Covid-19 pandemic on the cybersecurity in Italy}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Marco R. A. Bozzetti\inst{1}
\and
    Fausto Spoto\inst{2}
\and
   Luca Olivieri \inst{2}
}

% Institutes for affiliations are also joined by \and,
\institute{
  President AIPSI\\
  Italian Chapter ISSA
\and
   University of Verona,
   Italy\\
   \email{\{fausto.spoto,luca.olivieri\}@univr.it}\\
 }

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{M. Bozzetti, F.Spoto, L. Olivieri}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{The Covid-19 pandemic impacts on the cybersecurity in Italy}

\begin{document}

\maketitle

\begin{abstract}
%TODO
\end{abstract}

\section{Introduction}

OAD, the Digital Attacks Observatory in Italy, with the last 2020 edition \cite{OAD_report_2020} \footnote{The report is written in Italian, only 
the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters (147 A4 pages) and 9 Attachments (39 A4 pages)} reaches the 
twelfth year of consecutive surveys on cyberattacks in Italy, and, as in the previous years, with the precious collaboration of the Italian Postal 
and Telecommunications Police. OAD is the only independent online survey in Italy on intentional digital attacks on IT systems of companies and public
bodies operating in Italy. The OAD survey does not provide for a predefined set of respondents, but it allows potential interested parties full and 
free access to the online questionnaire, in a totally anonymous manner. Thanks to the number of responses collected and their balanced distribution 
between companies and public bodies of various sizes and belonging to various product sectors, the OAD survey provides a specific picture of the cyberattacks in Italy.

The OAD 2020 survey took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter of 2020, when in Italy the pandemic exploded.
The comparison of the data of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
clear indication of the Covid pandemic on the intentional cyber-attacks in Italy. This pandemic has been the trigger for a wide range of cyberattacks, mainly 
caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool  is well balanced for the size of the organizations, in terms
of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \footnote{According to the last ISTAT census}
of enterprises are SMEs, Small Medium Enterprises, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
very small organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
belong to structures with less than 250 employees, and of these 22.1\% under 10. The OAD questionaries refer for about half of the question to the cyber-attacks 
typologies and they main characteristics, if they have been relived. The other half of the questions refers to the technical and organizational security measures implemented
on the IT systems of the respondents. OAD therefore allows to contextualize the cyber-attacks with the cyber measures in production in each IT system considered. 
And in fact, in order to improve the motivation to complete the questionaries to all the potential anonymous respondents, OAD 2020 provided a macro qualitative evaluation
of the implemented measures, within the specific context of the company. The results of  this macro evaluation is shown in fig. 1, which shows that more than 2/3 of the respondents has a good level of cybersecurity.

%TODO: include fig 1

Fig. 2 compares the trend of the relived attacks from the beginning of the OAD survey in 2008: this comparison is only valid as a trend and not from a statistical point 
of view, due to the fact that the respondents pools year by year are different. Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). And in 2018 OAD 
reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.

%TODO: include fig 2

The dotted red line in the figure shows a wave shape: after a relative pick of attacks, the companies react hardening their cyber security, and the following year, normally, 
the relative pick in \% is lower that the previous one. After the biggest pick in 2018 as shown in fig.2,  2019 has a lower \%, and also for the first quarter 2020.
Only with the 2021 survey we will discover the \% for the whole 2020. The \%  in fig. 2 are lower than these expected.? We suggest to consider that:

\begin{itemize}
\item As shown in fig. 1, the majority of the respondents , in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
And therefore the IT Systems can properly react to the cyber attacks.

\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More r less is the same for the Public Administration (PA), 
the local ones in particular. And the number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
The hackers/crackers do not waist time for targeted attacks to very small realities.
\end{itemize}

Fig. 3 compares the cyber attacks relived in 2019 and in the first quarter of 2020 for the organization dimension. The black dotted line in the figure shows 
the “not” relieved attacks, and points out how it decrease from the very small to the very large organizations, in terms of numbers of employees.

%TODO: include fig 3

The figure confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
The comparison between the whole 2019 and the first quarter 2020 maintains the same logical trend. The only meaningful difference is for the frequent attacks relieved by the very
small organizations, up to 10 employees: 0\% in 2019 grows up to 22,2\% (see in fig. 3 the red line). Such a strong increase derives by the large utilization of e-baking and 
e-commerce services, with the related cyber attacks.

The impact of pandemic on the cyber security in Italy is also underlined by the data  provided by the Postal and Communications Police in fig. 4, that shows the situation for the
critical infrastructures, for the financial cyber crimes and for the cyber terrorism: and in particular for the financial cybercrimes. Fraudulent transactions in 2019 were blocked
for a total value of € 21.3 million and € 18 million were recovered; in the 1st quarter of 2020 fraudulent transactions for € 20.2 million were blocked, practically 
reaching in 4 months what was blocked in the 12 months of the previous year, and € 8 million recovered; an indicator of the growth in attacks on financial transactions and 
services due to the very large use of these online services caused by the mobility blocks imposed by the Covid-19 pandemic.

%TODO: include fig 4

The type of attack (what is attacked) most widespread in both the considered periods is the one aimed at access control systems (IAA, Identification Authentication Authorization), 
with 34\% in 2019 and 28.2\% in the 1st quarter of 2020, as shown in fig. 5. As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage
points among them, whose characteristics and impacts are detailed in the related paragraphs of the OAD 2020 report. Referring to the attack techniques (how to attack), all the seven
attacks families considered in the 2020 questionnaire are widely used in the various types of attacks, some times even simultaneously., as shown in fig. 6. The most widespread one,
as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
on the target system of the attack, with a 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.) with 34.6\% 
and the use of malicious code and scripts with 34.3%. The other considered techniques decrease with a few percentage points of difference between them. The impacts of the most 
critical attacks are analyzed for each attack type, as well as their possible motivations and the recovery times required by the most critical ones.

In the 2020 OAD survey, all these attacks’ characteristics vary for each type of attack, and the emerged results are described in the specific paragraphs dedicated to each attack 
type. At overall it emerges that:
\begin{itemize}
\item the impacts declared by the respondents are balanced between those irrelevant and / or easily resolvable with recovery in short time and at limited costs, and those very 
critical, which require expensive and long recovery actions, and that, in some cases, can cause business and customer loses; these two very different impact cases depend mainly 
on the security measures in place;

\item the motivations of the cyber-attacks are mainly of an economic nature, therefore done for fraud and blackmail: the widespread diffusion of ransomware in Italy is a clear 
confirmation of these motivations.
\end{itemize}

%TODO: include fig 5
%TODO: include fig 6
%TODO: include fig 7
%TODO: include fig 8

Fig. 8 shows the security measures considered in the 2020 questionnaire, subdivided in the technical ones, the organizational ones and  the management and governance ones. 
As pointed out in fig. 1, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the emerged information  show a relevant 
improvement in both technical and organizational measures, in comparison to the previous surveys. Few of the considered Italian Informatics Systems are based on Data Centers in 
Italy, most of the respondents’ companies have medium-small Informatic Systems partly on premise and partly outsourced, with an increasing use of cloud services.
The high number of attacks detected in 2018, and the privacy obligations (GDPR) have certainly contributed to strengthening cyber security measures, and a further improvement 
comes from the growing use of cloud services, where usually there are high standard security measures. Organizational measures for cybersecurity, historically lacking and neglected
in Italy, have improved in terms of defining roles and separation of duties, of organizational policies and procedures, and of incidents management. These measures often are lacking
in small organizations, and in general the cybersecurity awareness and competence are low, and mainly at the top level of the public and private organizations. In Italy, there is 
still a long way to go in terms of cybersecurity continual training and awareness. A formal and bureaucratic approach for the organizational procedures is still prevalent, which 
once defined, often do not find a possible concrete applicability, periodic updates and operational tests: an emblematic example is represented by the Disaster Recovery plans for 
which, several times, the required alternative ICT resources to be used are not forecasted and allocated and there also not provided periodic exercises and simulations of the 
possible disasters. Cybersecurity management tools still have limited diffusion among the 2020 OAD respondents, in particular the most advanced ones based on artificial intelligence
techniques. The use of IoT, of industrial automation and of robotics systems, and also of systems based on blockchain techniques find a low number of respondents involved, which 
also derives from the limited percentage of their product sectors that should be the more interested in these systems: the manufacturing sectors, the logistics, the research and 
development centers and labs, the Local Public Administrations for territorial control, and so on. For these topics, the data that emerged from the survey are considered only as 
specific cases that contribute to the general values on the cyberattacks, but which cannot be still considered of reference at the Italian level.

\section{Conclusion}

To conclude, the OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 
degrees related to the Covid-19 pandemic, in the first four months of 2020 the general spread of attacks is at similar values to those of 2019: to be verified with the next OAD 2021
if there will be changes to this trend.

The Italian reality, made up of a very large number of small and very small organizations, does not make our country one of the most attractive for cybercriminals, but cyber warfare
and massive attacks represent a growing and serious risk, as has already happened in part with the widespread of ransomware on computer systems whit a lack of or  with low 
cybersecurity  basic measures. OAD 2020 notes a clear improvement and strengthening of digital security measures, both technical and organizational, even if the most modern 
prevention, protection and management systems that use artificial intelligence techniques are still embryonic  in the respondents' basin.

The defense measures and techniques in use chase the increasingly sophisticated and smart evolution of attacks, but almost always late. The high density of vulnerabilities 
requires different approaches and new logics, with the aim of making all ICT systems interconnectable to the Internet intrinsically safe, by default and by design. But we are 
still a long way from this goal, and in order to decisively improve the concrete fight against continuous attacks and cybercrime, it is currently necessary to increase 
cybersecurity’ awareness and skills at all levels, an effective collaboration between the police at world level, and primarily a real and large usage of the professional ethics
 both of those involved (supply side) and both of those who decide (demand side) on cybersecurity.

\end{document}

### 2021-03-04 article.tex
Drafting an introduction and other minor changes

@@
 \documentclass{easychair}
-%\documentclass[EPiC]{easychair}
-%\documentclass[EPiCempty]{easychair}
-%\documentclass[debug]{easychair}
-%\documentclass[verbose]{easychair}
-%\documentclass[notimes]{easychair}
-%\documentclass[withtimes]{easychair}
-%\documentclass[a4paper]{easychair}
-%\documentclass[letterpaper]{easychair}
 
@@
 
-% use this if you have a long article and want to create an index
-% \usepackage{makeidx}
-
-% In order to save space or manage large tables or figures in a
-% landcape-like text, you can use the rotating and pdflscape
-% packages. Uncomment the desired from the below.
-%
-% \usepackage{rotating}
-% \usepackage{pdflscape}
-
-% Some of our commands for this guide.
-%
 \newcommand{\easychair}{\textsf{easychair}}
@@
 
-%\makeindex
+\title{Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy}
 
-%% Front Matter
-%%
-% Regular title as in the article class.
-%
-\title{The impacts of the beginning of the Covid-19 pandemic on the cybersecurity in Italy}
-
-% Authors are joined by \and. Their affiliations are given by \inst, which indexes
-% into the list defined using \institute
-%
 \author{
@@
 
-%  \authorrunning{} has to be set for the shorter version of the authors' names;
-% otherwise a warning will be rendered in the running heads. When processed by
-% EasyChair, this command is mandatory: a document without \authorrunning
-% will be rejected by EasyChair
-
 \authorrunning{M. Bozzetti, F.Spoto, L. Olivieri}
 
-% \titlerunning{} has to be set to either the main title or its shorter
-% version for the running heads. When processed by
-% EasyChair, this command is mandatory: a document without \titlerunning
-% will be rejected by EasyChair
-\titlerunning{The Covid-19 pandemic impacts on the cybersecurity in Italy}
+\titlerunning{Cybersecurity impacts at the beginning of Covid-19 pandemic in Italy}
 
@@
 
-OAD, the Digital Attacks Observatory in Italy, with the last 2020 edition \cite{OAD_report_2020} \footnote{The report is written in Italian, only 
-the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters (147 A4 pages) and 9 Attachments (39 A4 pages)} reaches the 
-twelfth year of consecutive surveys on cyberattacks in Italy, and, as in the previous years, with the precious collaboration of the Italian Postal 
-and Telecommunications Police. OAD is the only independent online survey in Italy on intentional digital attacks on IT systems of companies and public
-bodies operating in Italy. The OAD survey does not provide for a predefined set of respondents, but it allows potential interested parties full and 
-free access to the online questionnaire, in a totally anonymous manner. Thanks to the number of responses collected and their balanced distribution 
-between companies and public bodies of various sizes and belonging to various product sectors, the OAD survey provides a specific picture of the cyberattacks in Italy.
+The advent of the Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
+accelerating their digitalization efforts. As reported in \cite{Goasduff} by Gartner, a world's leading research and advisory company, before the pandemic crisis
+most organizations moved their digital strategies forward at a steady pace, but during the Covid-19 pandemic there has been a significant increase in the 
+development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security issues
+and the intensification of malicious cyber attacks around the world. In this scenario, Italy is an insteresting case of study because it is composed mostly from 
+Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies have a greater influence.
 
-The OAD 2020 survey took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter of 2020, when in Italy the pandemic exploded.
-The comparison of the data of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
+\section{Digital Attacks Observatory}
+
+The Digital Attacks Observatory (OAD)~\cite{OAD} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
+bodies operating and in addition drawn up with the precious collaboration of the Italian Postal and Telecommunications Police. The OAD survey does not provide 
+for a predefined set of respondents, but it allows potential interested parties full and free access to an online questionnaire, in a totally anonymous manner. 
+Thanks to the number of responses collected and their balanced distribution between companies and public bodies of various sizes and belonging to various product 
+sectors, the OAD survey provides a specific picture of the cyber-attacks in Italy.
+
+\section{Data comparison}
+
+OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
+(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{OAD_report_2020} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
+of 2020, when in Italy the pandemic exploded. The data comparison of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
 clear indication of the Covid pandemic on the intentional cyber-attacks in Italy. This pandemic has been the trigger for a wide range of cyberattacks, mainly 
@@
 even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool  is well balanced for the size of the organizations, in terms
-of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \footnote{According to the last ISTAT census}
-of enterprises are SMEs, Small Medium Enterprises, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
+of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\%
+\footnote{According to the last ISTAT census} %TODO: manca fonte/citazione
+of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
 very small organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
@@
 on the IT systems of the respondents. OAD therefore allows to contextualize the cyber-attacks with the cyber measures in production in each IT system considered. 
-And in fact, in order to improve the motivation to complete the questionaries to all the potential anonymous respondents, OAD 2020 provided a macro qualitative evaluation
-of the implemented measures, within the specific context of the company.
-The results of  this macro evaluation is shown in fig.
-1, which shows that more than 2/3 of the respondents has a good level of cybersecurity.
+In fact, OAD 2020 provided a macro qualitative evaluation of the implemented measures (within the specific context of the company), in order to improve the motivation to 
+complete the questionaries to all the potential anonymous respondents. The results of this macro evaluation is shown in fig. 1, which shows that more than $ 2/3 $ of the 
+respondents has a good level of cybersecurity.
 
-%TODO: include fig 1
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig1.png}
+		\caption{}
+		\label{fig:1}
+\end{figure}
 
@@
 of view, due to the fact that the respondents pools year by year are different. Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
-cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). And in 2018 OAD 
+cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). In 2018 OAD 
 reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.
 
-%TODO: include fig 2
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig2.png}
+		\caption{}
+		\label{fig:2}
+\end{figure}
 
@@
 \item As shown in fig. 1, the majority of the respondents , in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
-And therefore the IT Systems can properly react to the cyber attacks.
+Therefore the IT Systems can properly react to the cyber attacks.
 
 \item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More r less is the same for the Public Administration (PA), 
-the local ones in particular. And the number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
-The hackers/crackers do not waist time for targeted attacks to very small realities.
+the local ones in particular. The number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
+The malicious hackers/crackers do not waist time for targeted attacks to very small realities.
 \end{itemize}
@@
 
-%TODO: include fig 3
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig3.png}
+		\caption{}
+		\label{fig:3}
+\end{figure}
 
@@
 
-%TODO: include fig 4
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig4.png}
+		\caption{}
+		\label{fig:4}
+\end{figure}
 
@@
 
-%TODO: include fig 5
-%TODO: include fig 6
-%TODO: include fig 7
-%TODO: include fig 8
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig5.png}
+		\caption{}
+		\label{fig:5}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig6.png}
+		\caption{}
+		\label{fig:6}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig7.png}
+		\caption{}
+		\label{fig:7}
+\end{figure}
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig8.png}
+		\caption{}
+		\label{fig:8}
+\end{figure}
 
@@
 
-To conclude, the OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
+The OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
 spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 
@@
 
+\bibliographystyle{plain}
+\bibliography{biblio}
+
 \end{document}

### 2021-03-06 article.tex
Working on the paper

@@
 
+\providecommand{\keywords}[1]{\textbf{\textit{Index terms---}} #1}
+
 \title{Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy}
@@
 \author{
-Marco R. A. Bozzetti\inst{1}
+Marco R. A. Bozzetti\inst{1,2,3}
 \and
-    Fausto Spoto\inst{2}
+    Fausto Spoto\inst{4}
 \and
-   Luca Olivieri \inst{2}
+   Luca Olivieri \inst{4}
 }
@@
 \institute{
-  President AIPSI\\
-  Italian Chapter ISSA
+  Associazione Italiana Professionisti della Sicurezza Informatica (AIPSI), Milano, Italy
 \and
-   University of Verona,
-   Italy\\
+  Italian Information System Security Association International (ISSA)
+\and
+	Digital Attacks Observatory Team, Italy
+\and
+   University of Verona, Verona, Italy\\
    \email{\{fausto.spoto,luca.olivieri\}@univr.it}\\
@@
 %TODO
+
+  \keywords{Covid-19 impacts \and Cyberattacks \and Small Medium Enterprises \and Cybersecurity Culture.}
 \end{abstract}
@@
 The advent of the Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
-accelerating their digitalization efforts. As reported in \cite{Goasduff} by Gartner, a world's leading research and advisory company, before the pandemic crisis
+accelerating their digitalization efforts. As reported in \cite{Goasduff20} by Gartner, a world's leading research and advisory company, before the pandemic crisis
 most organizations moved their digital strategies forward at a steady pace, but during the Covid-19 pandemic there has been a significant increase in the 
-development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security issues
-and the intensification of malicious cyber attacks around the world. In this scenario, Italy is an insteresting case of study because it is composed mostly from 
-Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies have a greater influence.
+development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security 
+issues and the intensification of malicious cyber attacks around the world~\cite{HKICS20,PA21}. In this scenario, Italy is an interesting case of study because 
+it is composed mostly from Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies 
+have a greater influence.
 
-\section{Digital Attacks Observatory}
+This paper presents a window to understand how the cyber-security is faced up during COVID-19 pandemic in the italian panorama. 
+Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey containing relevant information about cyber security and italian organizations which will be used in subsequent
+sections as a reference to better understand Covid-19 impacts.
+A detailed discussions are in Sections~\ref{sec:DataDiscussion,sec:TypeAttacks}.
+Finally, Section~\ref{sec:Conclusion} concludes by underlying
+ the importance of cybersecurity culture focusing on Italian reality and SMEs.
 
-The Digital Attacks Observatory (OAD)~\cite{OAD} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
+\section{Digital Attacks Observatory}\label{sec:DigitalAttacksObservatory}
+
+The \textit{Digital Attacks Observatory} (OAD)~\cite{oadweb} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
 bodies operating and in addition drawn up with the precious collaboration of the Italian Postal and Telecommunications Police. The OAD survey does not provide 
@@
 Thanks to the number of responses collected and their balanced distribution between companies and public bodies of various sizes and belonging to various product 
-sectors, the OAD survey provides a specific picture of the cyber-attacks in Italy.
+sectors, the OAD survey provides a specific picture of the cyber attacks in Italy.
 
-\section{Data comparison}
+\subsection{Questionaries}
+
+The OAD questionaries refer for about half of the question to the cyber attacks typologies and they main characteristics, if they have been relived. 
+The other half of the questions refers to the technical and organizational security measures implemented on the IT systems of the respondents. 
+OAD therefore allows to contextualize the cyber attacks with the cyber measures in production in each IT system considered. 
+In fact, OAD provided a macro qualitative evaluation of the implemented measures (within the specific context of the company), in order to improve the motivation to 
+complete the questionaries to all the potential anonymous respondents.
+
+\subsection{Overview 2019-(1Q)2020}
 
 OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
-(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{OAD_report_2020} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
+(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{oad20} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
 of 2020, when in Italy the pandemic exploded. The data comparison of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
-clear indication of the Covid pandemic on the intentional cyber-attacks in Italy. This pandemic has been the trigger for a wide range of cyberattacks, mainly 
+clear indication of the Covid-19 pandemic on the intentional cyber attacks in Italy. This pandemic has been the trigger for a wide range of cyber attacks, mainly 
 caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
@@
 even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool  is well balanced for the size of the organizations, in terms
-of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\%
-\footnote{According to the last ISTAT census} %TODO: manca fonte/citazione
+of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat20} %TODO: manca fonte/citazione
 of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
-very small organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
-belong to structures with less than 250 employees, and of these 22.1\% under 10. The OAD questionaries refer for about half of the question to the cyber-attacks 
-typologies and they main characteristics, if they have been relived. The other half of the questions refers to the technical and organizational security measures implemented
-on the IT systems of the respondents. OAD therefore allows to contextualize the cyber-attacks with the cyber measures in production in each IT system considered. 
-In fact, OAD 2020 provided a macro qualitative evaluation of the implemented measures (within the specific context of the company), in order to improve the motivation to 
-complete the questionaries to all the potential anonymous respondents. The results of this macro evaluation is shown in fig. 1, which shows that more than $ 2/3 $ of the 
-respondents has a good level of cybersecurity.
+tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cyber security surveys: 57.5\% of the OAD 2020 respondents
+belong to structures with less than 250 employees, and of these 22.1\% under 10.
+
+\section{Data discussion}\label{sec:DataDiscussion}
+
+%TODO: scrivere un introduzione alla sezione
+
+The results of this macro evaluation are shown in fig. 1.
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig1.png}
-		\caption{}
+		\caption{The chart shows that more than $ 2/3 $ of the respondents has a good level of cybersecurity}
 		\label{fig:1}
@@
 
-Fig. 2 compares the trend of the relived attacks from the beginning of the OAD survey in 2008: this comparison is only valid as a trend and not from a statistical point 
-of view, due to the fact that the respondents pools year by year are different. Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
-cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). In 2018 OAD 
+Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
+cyber attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). As shown in Fig.2, in 2018 OAD 
 reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.
@@
 		\includegraphics[width=1\textwidth]{pictures/fig2.png}
-		\caption{}
+		\caption{Trend comparison of the relived attacks from the beginning of the OAD survey in 2008. This comparison is only valid as a trend and not from a statistical point 
+of view, due to the fact that the respondents pools year by year are different.
+The dotted red line shows a wave shape: after a relative pick of attacks, the companies react hardening their cyber security, and the following year, normally,
+the relative pick in \% is lower that the previous one.}
 		\label{fig:2}
@@
 
-The dotted red line in the figure shows a wave shape: after a relative pick of attacks, the companies react hardening their cyber security, and the following year, normally, 
-the relative pick in \% is lower that the previous one. After the biggest pick in 2018 as shown in fig.2,  2019 has a lower \%, and also for the first quarter 2020.
-Only with the 2021 survey we will discover the \% for the whole 2020. The \%  in fig. 2 are lower than these expected.? We suggest to consider that:
+After the biggest pick in 2018, 2019 has a lower \%, and also for the first quarter 2020. Only with the 2021 survey we will discover the \% for the whole 2020. The \%  in fig. 2
+are lower than these expected. We suggest to consider that:
 
 \begin{itemize}
-\item As shown in fig. 1, the majority of the respondents , in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
+\item As shown in fig. 1, the majority of the respondents, in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
 Therefore the IT Systems can properly react to the cyber attacks.
 
-\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More r less is the same for the Public Administration (PA), 
+\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More or less is the same for the Public Administration (PA), 
 the local ones in particular. The number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
-The malicious hackers/crackers do not waist time for targeted attacks to very small realities.
+The cybercriminals do not waist time for targeted attacks to very small realities.
 \end{itemize}
-
-Fig. 3 compares the cyber attacks relived in 2019 and in the first quarter of 2020 for the organization dimension. The black dotted line in the figure shows 
-the “not” relieved attacks, and points out how it decrease from the very small to the very large organizations, in terms of numbers of employees.
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig3.png}
-		\caption{}
+		\caption{Comparison between cyber attacks relived in 2019 and first quarter of 2020 for the organization dimension. The black dotted line in the figure shows 
+the “not” relieved attacks, and points out how it decrease from the tiny to the big organizations, in terms of numbers of employees. }
 		\label{fig:3}
@@
 
-The figure confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
+The figure Fig. 3 confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
 The comparison between the whole 2019 and the first quarter 2020 maintains the same logical trend. The only meaningful difference is for the frequent attacks relieved by the very
@@
 
+\section{Type of attacks}\label{sec:TypeAttacks}
+
 The type of attack (what is attacked) most widespread in both the considered periods is the one aimed at access control systems (IAA, Identification Authentication Authorization), 
@@
 
-\item the motivations of the cyber-attacks are mainly of an economic nature, therefore done for fraud and blackmail: the widespread diffusion of ransomware in Italy is a clear 
+\item the motivations of the cyber attacks are mainly of an economic nature, therefore done for fraud and blackmail: the widespread diffusion of ransomware in Italy is a clear 
 confirmation of these motivations.
@@
 Italy, most of the respondents’ companies have medium-small Informatic Systems partly on premise and partly outsourced, with an increasing use of cloud services.
-The high number of attacks detected in 2018, and the privacy obligations (GDPR) have certainly contributed to strengthening cyber security measures, and a further improvement 
+The high number of attacks detected in 2018, and the privacy obligations, related to General Data Protection Regulation (GDPR), have certainly contributed to strengthening cyber security measures, and a further improvement
 comes from the growing use of cloud services, where usually there are high standard security measures. Organizational measures for cybersecurity, historically lacking and neglected
@@
 
-\section{Conclusion}
+\section{Conclusion}\label{sec:Conclusion}
 
@@
 
-The Italian reality, made up of a very large number of small and very small organizations, does not make our country one of the most attractive for cybercriminals, but cyber warfare
+The Italian reality, made up of a very large number of small and tiny organizations, does not make our country one of the most attractive for cybercriminals, but cyber warfare
 and massive attacks represent a growing and serious risk, as has already happened in part with the widespread of ransomware on computer systems whit a lack of or  with low 

### 2021-03-06 article.tex
Working on draft

@@
 
-\providecommand{\keywords}[1]{\textbf{\textit{Index terms---}} #1}
+\usepackage{eurosym}
+\usepackage{graphicx}
 
@@
 \and
-	Digital Attacks Observatory Team, Italy
+	Digital Attacks Observatory Team (OAD), Italy
 \and
@@
 
-  \keywords{Covid-19 impacts \and Cyberattacks \and Small Medium Enterprises \and Cybersecurity Culture.}
+  %\keywords{Covid-19 impacts \and Cyberattacks \and Small Medium Enterprises \and Cybersecurity Culture.}
 \end{abstract}
@@
 
-This paper presents a window to understand how the cyber-security is faced up during COVID-19 pandemic in the italian panorama. 
-Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey containing relevant information about cyber security and italian organizations which will be used in subsequent
-sections as a reference to better understand Covid-19 impacts.
-A detailed discussions are in Sections~\ref{sec:DataDiscussion,sec:TypeAttacks}.
-Finally, Section~\ref{sec:Conclusion} concludes by underlying
- the importance of cybersecurity culture focusing on Italian reality and SMEs.
+This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
+Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey containing relevant information about cyber security and 
+italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of results are in 
+Sections~\ref{sec:DataDiscussion,sec:TypeAttacks}. Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of cybersecurity culture focusing 
+on Italian reality and SMEs.
 
@@
 complete the questionaries to all the potential anonymous respondents. 
+In particular the security measures considered in the 2020 questionnaire are subdivided in technical, organizational, management and governance ones as follow:
+\begin{enumerate}
+\small
+	\item \emph{Technical measures}
+	\begin{itemize}
+		\item Overall architecture of digital security measures, integrated with the entire IT system architecture
+		\item Physical countermeasures
+		\item Identification, Authentication, and Authorization (IAA) measures
+		\item Cyber security measures at local network and geographic level
+		\item Measures for logical protection of each single Information and Communications Technology (ICT) system
+		\item Application and software protection
+		\item Data protection
+	\end{itemize}
+	\item \emph{Organizational measures}
+	\begin{itemize}
+		\item Organizational structure, roles, skills, certifications
+		\item Organizational policies and procedures
+		\item Contracts, agreements and digital security clauses with third parties
+		\item Awareness and culture of digital security at all levels
+		\item Auditing
+	\end{itemize}
+	\item \emph{Management and governance measures}
+	\begin{itemize}
+		\item Digital security control, monitoring and management systems
+		\item Disaster Recovery (DR) plan
+	\end{itemize}
+\end{enumerate}
 
@@
 
-%TODO: scrivere un introduzione alla sezione
-
-The results of this macro evaluation are shown in fig. 1.
-
 \begin{figure}
@@
 		\includegraphics[width=1\textwidth]{pictures/fig1.png}
-		\caption{The chart shows that more than $ 2/3 $ of the respondents has a good level of cybersecurity}
+		\caption{Extract of OAD 2020 \cite{oad20} on macro evaluation of security level in the information systems in Italy.
+The chart shows that more than $ 2/3 $ of the respondents has a good level of cybersecurity.}
 		\label{fig:1}
@@
 
-Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
-cyber attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). As shown in Fig.2, in 2018 OAD 
-reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.
+The data collected in the OAD 2020 showed at a high level for the most part a good perception by the respondents of their IT security (Fig.~\ref{fig:1}).
+This information can be explainable from the fact that:
+
+\begin{itemize}
+\item About 30\% of respondents are in the ICT sector, therefore the IT Systems already known the cybersecurity culture and can properly react to the cyber attacks, preventing 
+attacks and minimizing risks.
+
+\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. Specifically in the 2020 OAD survey, the respondents are 57,7\% for 
+SME/small P, and of these 22,1\% are under 10 employees. In general, cybercriminals do not waste time in attacks targeting small businesses, to obtain only a marginal profit, 
+with the risk of being caught by the police.
+\end{itemize}
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig2.png}
-		\caption{Trend comparison of the relived attacks from the beginning of the OAD survey in 2008. This comparison is only valid as a trend and not from a statistical point 
+		\caption{Trend comparison of the relived attacks from the beginning of the OAD survey in 2008~\cite{oad20}. This comparison is only valid as a trend and not from a statistical point 
 of view, due to the fact that the respondents pools year by year are different.
@@
 
-After the biggest pick in 2018, 2019 has a lower \%, and also for the first quarter 2020. Only with the 2021 survey we will discover the \% for the whole 2020. The \%  in fig. 2
-are lower than these expected. We suggest to consider that:
-
-\begin{itemize}
-\item As shown in fig. 1, the majority of the respondents, in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
-Therefore the IT Systems can properly react to the cyber attacks.
-
-\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More or less is the same for the Public Administration (PA), 
-the local ones in particular. The number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
-The cybercriminals do not waist time for targeted attacks to very small realities.
-\end{itemize}
+Analyzing the trends about cyber attaks since 2007 to 2015, the chart in Fig.~\ref{fig:2} reveals in average that about 40\% of the respondents that relived cyber attacks 
+(first red line). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line). 
+In 2018 OAD reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that percentage of the not relived. After the biggest pick 
+in 2018, 2019 has a lower percentage, and also for the first quarter 2020. Only with the 2021 survey we will discover the percentages for the whole 2020. 
+This view over several years of the interviewees does not show significant changes especially compared to the pandemic period but only a slight increase in cyber attacks 
+in recent years not necessarily attributable to the pandemic crisis.
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig3.png}
-		\caption{Comparison between cyber attacks relived in 2019 and first quarter of 2020 for the organization dimension. The black dotted line in the figure shows 
+		\caption{Comparison between cyber attacks relived in 2019 and first quarter of 2020 for the organization dimension~\cite{oad20}. The black dotted line in the figure shows 
 the “not” relieved attacks, and points out how it decrease from the tiny to the big organizations, in terms of numbers of employees. }
@@
 
-The figure Fig. 3 confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
+Fig.~\ref{fig:3} confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
 The comparison between the whole 2019 and the first quarter 2020 maintains the same logical trend. The only meaningful difference is for the frequent attacks relieved by the very
-small organizations, up to 10 employees: 0\% in 2019 grows up to 22,2\% (see in fig. 3 the red line). Such a strong increase derives by the large utilization of e-baking and 
+small organizations, up to 10 employees: 0\% in 2019 grows up to 22,2\% (see in Fig.~\ref{fig:3} the red line). Such a strong increase derives by the large utilization of e-baking and 
 e-commerce services, with the related cyber attacks.
 
-The impact of pandemic on the cyber security in Italy is also underlined by the data provided by the Postal and Communications Police in fig. 4, that shows the situation for the
-critical infrastructures, for the financial cyber crimes and for the cyber terrorism: and in particular for the financial cybercrimes. Fraudulent transactions in 2019 were blocked
-for a total value of € 21.3 million and € 18 million were recovered; in the 1st quarter of 2020 fraudulent transactions for € 20.2 million were blocked, practically 
-reaching in 4 months what was blocked in the 12 months of the previous year, and € 8 million recovered; an indicator of the growth in attacks on financial transactions and 
-services due to the very large use of these online services caused by the mobility blocks imposed by the Covid-19 pandemic.
+The impact of pandemic on the cyber security in Italy is also highlighted by the data provided by the Postal and Communications Police that shows the situation for the
+critical infrastructures (Table.\ref{tab:table1}), for the financial cyber crimes (Table.\ref{tab:table2}) and for the cyber terrorism (Table.\ref{tab:table3}). 
+Fraudulent transactions in 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the 1st quarter of 
+2020 fraudulent transactions for \euro 20.2 million were blocked, practically reaching in 4 months what was blocked in the 12 months of the previous year, and \euro 8 million 
+recovered; an indicator of the growth in attacks on financial transactions and services due to the very large use of these online services caused by the mobility blocks imposed 
+by the Covid-19 pandemic.
 
-\begin{figure}
-	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig4.png}
-		\caption{}
-		\label{fig:4}
-\end{figure}
+\begin{table}[h]
+  \begin{center}
+  \resizebox{\textwidth}{!}{
+	\begin{tabular}{|r|r|r|r|r|r|}									
+		\hline									
+		Critical structure protection	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018 &	1 Jan - 31 Dec 2017 &	1 Jan - 31 Dec 2016\\
+		\hline									
+		Relevant attacks	&	282	&	1.181	&	459	&	1.032 & 844	\\
+		Issued alerts	&	24.824	&	82.484	&	80.777	&	31.524 & 6.721	\\
+		Initiated investigations 	&	34	&	155	&	74	&	72 & 70	\\
+		People arrested	&	0	&	3	&	1	&	3	& 3\\
+		People reported	&	0	&	117	&	14	&	1.316 &	1.226	\\
+		Perquisitions	&	n.d.	&	n.d.	&	n.d.	&	73	& 58 \\
+		\hline	
+		Initiated investigations on detected attacks	&	12,05\%	&	13,12\%	&	16,12\%	&	6,98\%	& 8,29\% \\
+		\hline	
+		People arrested out of	reported &	0\%	&	2,5\%	&	7,14\%	&	0,23\%	&	0,24\%\\						
+		\hline	
+	\end{tabular}
+	}
+    \end{center}
+
+	\caption{Data of critical structure protection collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\label{tab:table1}									
+\end{table}
+
+\begin{table}[h]
+  \begin{center}
+  \resizebox{\textwidth}{!}{
+	\begin{tabular}{|r|r|r|r|r|r|}									
+		\hline									
+		Financial Cyber Crime	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018 &	1 Jan - 31 Dec 2017 &	1 Jan - 31 Dec 2016\\
+		\hline									
+		Blocked fraudulent transactions	&	\textbf{\euro 20.200.000,00}	&	\textbf{\euro 21.333.990,00}	&	\euro 38.400.000,00	& \euro 20.839.576,00 & \euro 16.050.812,50 	\\
+		Recovered amounts 	&	\euro 8.700.000,00	&	\euro 18.000.000,00	&	\euro 9.000.000,00	&	\euro 862.000,00 & n.d.	\\
+		People arrested	&	n.d.	&	n.d.	&	n.d.	&	25	& 25\\
+		People reported	&	n.d.	&	n.d.	&	n.d.	&	2.851 & 3.772	\\
+		\hline	
+		Recovered amounts out of fraud	&	12,05\%	&	13,12\%	&	16,12\%	&	6,98\%	& 8,29\% \\
+		\hline	
+	\end{tabular}
+	}
+    \end{center}
+	\caption{Data of critical structure protection collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\label{tab:table2}									
+\end{table}
+
+\begin{table}[h]
+  \begin{center}
+	\begin{tabular}{|r|r|r|r|}									
+		\hline									
+		Cyber terrorism	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018\\
+		\hline									
+		Monitored web space	&	\euro 11.962	&	\euro 36.377	&	\euro 36.000\\
+		Removed contents 	&	n.d.	&	n.d.	&	250\\
+		\hline	
+	\end{tabular}
+        \end{center}
+	\caption{Data of cyber terrorism collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\label{tab:table3}									
+\end{table}
 
@@
 The type of attack (what is attacked) most widespread in both the considered periods is the one aimed at access control systems (IAA, Identification Authentication Authorization), 
-with 34\% in 2019 and 28.2\% in the 1st quarter of 2020, as shown in fig. 5. As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage
+with 34\% in 2019 and 28.2\% in the 1st quarter of 2020, as shown in Fig.~\ref{fig:5}. As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage
 points among them, whose characteristics and impacts are detailed in the related paragraphs of the OAD 2020 report. Referring to the attack techniques (how to attack), all the seven
-attacks families considered in the 2020 questionnaire are widely used in the various types of attacks, some times even simultaneously., as shown in fig. 6. The most widespread one,
+attacks families considered in the 2020 questionnaire are widely used in the various types of attacks, some times even simultaneously., as shown in Fig.~\ref{fig:6}. The most widespread one,
 as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
 on the target system of the attack, with a 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.) with 34.6\% 
-and the use of malicious code and scripts with 34.3%. The other considered techniques decrease with a few percentage points of difference between them. The impacts of the most 
+and the use of malicious code and scripts with 34.3%. The other considered technolog decrease with a few percentage points of difference between them. The impacts of the most 
 critical attacks are analyzed for each attack type, as well as their possible motivations and the recovery times required by the most critical ones.
@@
 \end{figure}
-\begin{figure}
-	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig7.png}
-		\caption{}
-		\label{fig:7}
-\end{figure}
-\begin{figure}
-	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig8.png}
-		\caption{}
-		\label{fig:8}
-\end{figure}
 
-Fig. 8 shows the security measures considered in the 2020 questionnaire, subdivided in the technical ones, the organizational ones and  the management and governance ones. 
-As pointed out in fig. 1, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the emerged information show a relevant 
+As pointed out in Fig.~\ref{fig:1}, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the emerged information show a relevant 
 improvement in both technical and organizational measures, in comparison to the previous surveys. Few of the considered Italian Informatics Systems are based on Data Centers in 
@@
 possible disasters. Cybersecurity management tools still have limited diffusion among the 2020 OAD respondents, in particular the most advanced ones based on artificial intelligence
-techniques. The use of IoT, of industrial automation and of robotics systems, and also of systems based on blockchain techniques find a low number of respondents involved, which 
+techniques. The use of IoT, of industrial automation and of robotics systems, and also of systems based on blockchain technologies find a low number of respondents involved, which 
 also derives from the limited percentage of their product sectors that should be the more interested in these systems: the manufacturing sectors, the logistics, the research and 
@@
 if there will be changes to this trend.
-
 The Italian reality, made up of a very large number of small and tiny organizations, does not make our country one of the most attractive for cybercriminals, but cyber warfare
@@
 cybersecurity  basic measures. OAD 2020 notes a clear improvement and strengthening of digital security measures, both technical and organizational, even if the most modern 
-prevention, protection and management systems that use artificial intelligence techniques are still embryonic  in the respondents' basin.
-
-The defense measures and techniques in use chase the increasingly sophisticated and smart evolution of attacks, but almost always late. The high density of vulnerabilities 
+prevention, protection and management systems that use artificial intelligence techniques are still embryonic  in the respondents' basin. The defense measures and techniques 
+in use chase the increasingly sophisticated and smart evolution of attacks, but almost always late. The high density of vulnerabilities 
 requires different approaches and new logics, with the aim of making all ICT systems interconnectable to the Internet intrinsically safe, by default and by design. But we are 

### 2021-03-06 article.tex
Working on draft

@@
 \and
-	Digital Attacks Observatory Team (OAD), Italy
+	Digital Attacks Observatory (OAD) Team, Italy
 \and
@@
 
-The advent of the Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
+The advent of Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
 accelerating their digitalization efforts. As reported in \cite{Goasduff20} by Gartner, a world's leading research and advisory company, before the pandemic crisis
@@
 it is composed mostly from Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies 
-have a greater influence.
-
-This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
+have a greater influence. This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
 Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey containing relevant information about cyber security and 
-italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of results are in 
-Sections~\ref{sec:DataDiscussion,sec:TypeAttacks}. Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of cybersecurity culture focusing 
+italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of collected data are in 
+Sections~\ref{sec:DataDiscussion}, \ref{sec:TypeAttacks}. Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of cybersecurity culture focusing 
 on Italian reality and SMEs.
@@
 
-The OAD questionaries refer for about half of the question to the cyber attacks typologies and they main characteristics, if they have been relived. 
+The OAD questionaries refer for about half of the question to cyber attacks typologies and they main characteristics, if they have been relived. 
 The other half of the questions refers to the technical and organizational security measures implemented on the IT systems of the respondents. 
@@
 (147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{oad20} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
-of 2020, when in Italy the pandemic exploded. The data comparison of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
+of 2020, when in Italy the pandemic exploded. The data comparison of 2019 and the first quarter 2020, provided by the same respondents pool, has a statistical relevance and provides a 
 clear indication of the Covid-19 pandemic on the intentional cyber attacks in Italy. This pandemic has been the trigger for a wide range of cyber attacks, mainly 
@@
 even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
-of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat20} %TODO: manca fonte/citazione
+of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat20} %TODO: manca fonte/citazione
 of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
@@
 		\caption{Comparison between cyber attacks relived in 2019 and first quarter of 2020 for the organization dimension~\cite{oad20}. The black dotted line in the figure shows 
-the “not” relieved attacks, and points out how it decrease from the tiny to the big organizations, in terms of numbers of employees. }
+the “not” relieved attacks, and points out how it decrease from the tiny to the big organizations, in terms of employees. }
 		\label{fig:3}
@@
   \begin{center}
-  \resizebox{\textwidth}{!}{
+  \resizebox{1.15\textwidth}{!}{
 	\begin{tabular}{|r|r|r|r|r|r|}									
@@
 
-The type of attack (what is attacked) most widespread in both the considered periods is the one aimed at access control systems (IAA, Identification Authentication Authorization), 
-with 34\% in 2019 and 28.2\% in the 1st quarter of 2020, as shown in Fig.~\ref{fig:5}. As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage
-points among them, whose characteristics and impacts are detailed in the related paragraphs of the OAD 2020 report. Referring to the attack techniques (how to attack), all the seven
-attacks families considered in the 2020 questionnaire are widely used in the various types of attacks, some times even simultaneously., as shown in Fig.~\ref{fig:6}. The most widespread one,
+In the pandemic period between 2019 and 1st quarter of 2020, the attacks on IAA that aimed at access control systems are most widespread with 34\% in 2019 and 28.2\% (Fig.~\ref{fig:5}).
+As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage points among them, whose characteristics and impacts are detailed in the related 
+paragraphs of the OAD 2020 report. Regarding attack vectors and techniques, all the seven attacks families considered in the 2020 questionnaire are widely used in the various 
+types of attacks, some times even simultaneously., as shown in Fig.~\ref{fig:6}. The most widespread one,
 as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
 on the target system of the attack, with a 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.) with 34.6\% 
-and the use of malicious code and scripts with 34.3%. The other considered technolog decrease with a few percentage points of difference between them. The impacts of the most 
+and the use of malicious code and scripts with 34.3\%. The other considered technolog decrease with a few percentage points of difference between them. The impacts of the most 
 critical attacks are analyzed for each attack type, as well as their possible motivations and the recovery times required by the most critical ones.
-
 In the 2020 OAD survey, all these attacks’ characteristics vary for each type of attack, and the emerged results are described in the specific paragraphs dedicated to each attack 
@@
 \begin{itemize}
-\item the impacts declared by the respondents are balanced between those irrelevant and / or easily resolvable with recovery in short time and at limited costs, and those very 
+\item the impacts declared by the respondents are balanced between those irrelevant and/or easily resolvable with recovery in short time and at limited costs, and those very 
 critical, which require expensive and long recovery actions, and that, in some cases, can cause business and customer loses; these two very different impact cases depend mainly 
@@
 possible disasters. Cybersecurity management tools still have limited diffusion among the 2020 OAD respondents, in particular the most advanced ones based on artificial intelligence
-techniques. The use of IoT, of industrial automation and of robotics systems, and also of systems based on blockchain technologies find a low number of respondents involved, which 
-also derives from the limited percentage of their product sectors that should be the more interested in these systems: the manufacturing sectors, the logistics, the research and 
+techniques.
+Although the use of IoT \cite{JK21}, industrial automation, robotics systems \cite{SGLMDXHKCZT21}, and systems based on blockchain technologies \cite{RMFF21} are expanding and can allow to combat the Covid-19 pandemic, in
+the OAD 2020, we can find only a low number of respondents involved, which also derives from the limited percentage of their product sectors that should be the more interested in these systems: the manufacturing sectors, the logistics, the research and
 development centers and labs, the Local Public Administrations for territorial control, and so on. For these topics, the data that emerged from the survey are considered only as 

### 2021-03-06 article.tex
added istat reference

@@
 even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
-of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat20} %TODO: manca fonte/citazione
+of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat21}
 of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 

### 2021-03-08 article.tex
Minor fix

@@
 issues and the intensification of malicious cyber attacks around the world~\cite{HKICS20,PA21}. In this scenario, Italy is an interesting case of study because 
-it is composed mostly from Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies 
-have a greater influence. This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
-Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey containing relevant information about cyber security and 
-italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of collected data are in 
+it is composed mostly from Small Medium Enterprises (SMEs) of different sectors and therefore can highlight specific aspects instead of the various global views 
+in which large companies have a greater influence. This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
+Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey where we have collected relevant information about cyber security and 
+italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of collected data and trends are in 
 Sections~\ref{sec:DataDiscussion}, \ref{sec:TypeAttacks}. Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of cybersecurity culture focusing 
-on Italian reality and SMEs.
+on Italian reality.
 
@@
 Thanks to the number of responses collected and their balanced distribution between companies and public bodies of various sizes and belonging to various product 
-sectors, the OAD survey provides a specific picture of the cyber attacks in Italy.
+sectors, we can provide a specific picture of the cyber attacks in Italy.  The last OAD 2020  reaches the twelfth year of consecutive surveys on cyberattacks in Italy.
 
@@
 even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
-of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \cite{istat21}
+of employees, between those below 250 and the largest ones. It must be taken into account, as reported by ISTAT~\cite{istat21}, that in Italy 99.91\%
 of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
@@
 
-\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. Specifically in the 2020 OAD survey, the respondents are 57,7\% for 
-SME/small P, and of these 22,1\% are under 10 employees. In general, cybercriminals do not waste time in attacks targeting small businesses, to obtain only a marginal profit, 
-with the risk of being caught by the police.
+\item In general, cybercriminals do not waste time in attacks targeting small businesses, to obtain only a marginal profit, 
+with the risk of being caught by the police.  As confirmed by trends of Fig.~\ref{fig:3}, the attackers try to hit the big structures,
+ because with these it is more probable the realization of gains, even if illegal.
 \end{itemize}
@@
 
-Fig.~\ref{fig:3} confirms what already underlined before: the attackers try to hit the big structures, because with these it is more probable the realization of gains, even if illegal.
-The comparison between the whole 2019 and the first quarter 2020 maintains the same logical trend. The only meaningful difference is for the frequent attacks relieved by the very
+As already highlighted above, the comparison between the whole 2019 and the first quarter 2020 in Fig.~\ref{fig:3} maintains the same logical trend.
+The only meaningful difference is for the frequent attacks relieved by the very
 small organizations, up to 10 employees: 0\% in 2019 grows up to 22,2\% (see in Fig.~\ref{fig:3} the red line). Such a strong increase derives by the large utilization of e-baking and 
@@
 
-\section{Type of attacks}\label{sec:TypeAttacks}
+\section{Cyber attack trends}\label{sec:TypeAttacks}
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig5.png}
-		\caption{}
+		\caption{Diffusion of digital attacks on respondents' computer systems based on the type of attack between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
 		\label{fig:5}
@@
 	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig6.png}
-		\caption{}
+		\includegraphics[width=0.90\textwidth]{pictures/fig6.png}
+		\caption{Average diffusion (percentage) of the attack techniques detected in the most serious attacks between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
 		\label{fig:6}
@@
 
-As pointed out in Fig.~\ref{fig:1}, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the emerged information show a relevant 
+As pointed out in Fig.~\ref{fig:1}, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the collected information show a relevant 
 improvement in both technical and organizational measures, in comparison to the previous surveys. Few of the considered Italian Informatics Systems are based on Data Centers in 
@@
 
-The OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
+The OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of an high complexity and sophistication, with a slightly increasing 
 spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 

### 2021-03-08 article.tex
Added enisa cite

@@
 most organizations moved their digital strategies forward at a steady pace, but during the Covid-19 pandemic there has been a significant increase in the 
-development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security 
+development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cybersecurity 
 issues and the intensification of malicious cyber attacks around the world~\cite{HKICS20,PA21}. In this scenario, Italy is an interesting case of study because 
@@
 in which large companies have a greater influence. This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
-Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey where we have collected relevant information about cyber security and 
+Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey where we have collected relevant information about cybersecurity and 
 italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of collected data and trends are in 
@@
 		\item Identification, Authentication, and Authorization (IAA) measures
-		\item Cyber security measures at local network and geographic level
+		\item Cybersecurity measures at local network and geographic level
 		\item Measures for logical protection of each single Information and Communications Technology (ICT) system
@@
 of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
-tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cyber security surveys: 57.5\% of the OAD 2020 respondents
+tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
 belong to structures with less than 250 employees, and of these 22.1\% under 10.
@@
 of view, due to the fact that the respondents pools year by year are different.
-The dotted red line shows a wave shape: after a relative pick of attacks, the companies react hardening their cyber security, and the following year, normally,
+The dotted red line shows a wave shape: after a relative pick of attacks, the companies react hardening their cybersecurity, and the following year, normally,
 the relative pick in \% is lower that the previous one.}
@@
 
-The impact of pandemic on the cyber security in Italy is also highlighted by the data provided by the Postal and Communications Police that shows the situation for the
+The impact of pandemic on the cybersecurity in Italy is also highlighted by the data provided by the Postal and Communications Police that shows the situation for the
 critical infrastructures (Table.\ref{tab:table1}), for the financial cyber crimes (Table.\ref{tab:table2}) and for the cyber terrorism (Table.\ref{tab:table3}). 
@@
 paragraphs of the OAD 2020 report. Regarding attack vectors and techniques, all the seven attacks families considered in the 2020 questionnaire are widely used in the various 
-types of attacks, some times even simultaneously., as shown in Fig.~\ref{fig:6}. The most widespread one,
-as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
+types of attacks, some times even simultaneously, as shown in Fig.~\ref{fig:6}.
+The trend of Fig.~\ref{fig:6} also mirrors for the most part the european threads recorded by European Network and Information Security Agency (ENISA)
+in the same period~\cite{enisa20}.
+The most widespread one, as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
 on the target system of the attack, with a 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.) with 34.6\% 
-and the use of malicious code and scripts with 34.3\%. The other considered technolog decrease with a few percentage points of difference between them. The impacts of the most 
+and the use of malicious code and scripts with 34.3\%. The other considered technologies decrease with a few percentage points of difference between them. The impacts of the most 
 critical attacks are analyzed for each attack type, as well as their possible motivations and the recovery times required by the most critical ones.
@@
 \end{itemize}
+
+In particular, the European Union Agency For Cybersecurity(ENISA)
 
@@
 Italy, most of the respondents’ companies have medium-small Informatic Systems partly on premise and partly outsourced, with an increasing use of cloud services.
-The high number of attacks detected in 2018, and the privacy obligations, related to General Data Protection Regulation (GDPR), have certainly contributed to strengthening cyber security measures, and a further improvement
+The high number of attacks detected in 2018, and the privacy obligations, related to General Data Protection Regulation (GDPR), have certainly contributed to strengthening cybersecurity measures, and a further improvement
 comes from the growing use of cloud services, where usually there are high standard security measures. Organizational measures for cybersecurity, historically lacking and neglected

### 2021-03-08 article.tex
Minor fix
Minor changes

@@
 
-In the pandemic period between 2019 and 1st quarter of 2020, the attacks on IAA that aimed at access control systems are most widespread with 34\% in 2019 and 28.2\% (Fig.~\ref{fig:5}).
+In the pandemic period between 2019 and 1st quarter of 2020,  the nature of attacks and their vectors were quite diversified.
+The attacks on IAA that aimed at access control systems are most widespread with 34\% in 2019 and 28.2\% (Fig.~\ref{fig:5}).
+
+\begin{figure}
+	\centering
+		\includegraphics[width=1\textwidth]{pictures/fig5.png}
+		\caption{Diffusion of digital attacks on respondents' computer systems based on the type of attack between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
+		\label{fig:5}
+\end{figure}
+
 As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage points among them, whose characteristics and impacts are detailed in the related 
@@
 types of attacks, some times even simultaneously, as shown in Fig.~\ref{fig:6}.
-The trend of Fig.~\ref{fig:6} also mirrors for the most part the european threads recorded by European Network and Information Security Agency (ENISA)
-in the same period~\cite{enisa20}.
+The trend of Fig.~\ref{fig:6} is also in line with for the most part the european threads recorded in ~\cite{enisa20} by European Network and Information Security Agency (ENISA).
 The most widespread one, as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
@@
 
-In particular, the European Union Agency For Cybersecurity(ENISA)
-
-\begin{figure}
-	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig5.png}
-		\caption{Diffusion of digital attacks on respondents' computer systems based on the type of attack between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
-		\label{fig:5}
-\end{figure}
 \begin{figure}

### 2021-03-08 article.tex
Clean-up of the paper

--- 2021-03-08 article.tex
@@
 
-\title{Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy}
+\title{Cybersecurity Impacts of the Covid-19 Pandemic in Italy}
 
 \author{
-Marco R. A. Bozzetti\inst{1,2,3}
+  Marco R.A.\ Bozzetti\inst{1,2,3}
+  \and
+  Luca Olivieri\inst{4}
   \and
   Fausto Spoto\inst{4}
-\and
-   Luca Olivieri \inst{4}
 }
@@
    University of Verona, Verona, Italy\\
-   \email{\{fausto.spoto,luca.olivieri\}@univr.it}\\
+   \email{\{luca.olivieri,fausto.spoto\}@univr.it}\\
  }
 
-\authorrunning{M. Bozzetti, F.Spoto, L. Olivieri}
+\authorrunning{M.\ Bozzetti, L.\ Olivieri, F.\ Spoto}
 
@@
 \begin{abstract}
-%TODO
+  The Covid-19 pandemic has pushed companies to the extensive
+  use of digital services, to implement home working and provide
+  online services to people in lockdown. As a consequence, it is interesting
+  to study how this has affected the number, kind and distribution of cybersecurity
+  attacks. This paper gives an empirical evaluation of the cybersecurity attacks
+  at the beginning of the Covid-19 pandemic in Italy, based on data collected from
+  the questionnaires of the annual Digital Attacks Observatory. It shows that the overall
+  number of attacks has not increased, but attacks have affected smaller companies
+  than before. This can be explained with the fact that the Italian
+  industrial scenario is mostly populated by small and medium enterprises,
+  that have been obliged to a quick reconvertion of their IT systems and typically
+  lack the necessary cybersecurity culture.
 
-  %\keywords{Covid-19 impacts \and Cyberattacks \and Small Medium Enterprises \and Cybersecurity Culture.}
+  \mbox{}\\
+  \textbf{Keywords:} cyberattack, cybersecurity, Covid-19, SME.
 \end{abstract}
@@
 
-The advent of Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
-accelerating their digitalization efforts. As reported in \cite{Goasduff20} by Gartner, a world's leading research and advisory company, before the pandemic crisis
-most organizations moved their digital strategies forward at a steady pace, but during the Covid-19 pandemic there has been a significant increase in the 
-development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security 
-issues and the intensification of malicious cyber attacks around the world~\cite{HKICS20,PA21}. In this scenario, Italy is an interesting case of study because 
-it is composed mostly from Small Medium Enterprises (SMEs) of different sectors and therefore can highlight specific aspects instead of the various global views 
-in which large companies have a greater influence. This paper presents a window of italian panorama to understand how the cyber-security is faced up during COVID-19 pandemic. 
-Section~\ref{sec:DigitalAttacksObservatory} presents \textit{Digital Attacks Observatory}, a survey where we have collected relevant information about cyber security and 
-italian organizations which will be used in subsequent sections as a reference to better understand Covid-19 impacts. A detailed discussions of collected data and trends are in 
-Sections~\ref{sec:DataDiscussion}, \ref{sec:TypeAttacks}. Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of cybersecurity culture focusing 
-on Italian reality.
+The Covid-19 pandemic and consequent lockdowns have obliged
+companies to face new challenges such as smart working, remote work and digitalization,
+accelerating all previous efforts in that direction.
+As reported by Gartner~\cite{Goasduff20}, most organizations were already moving
+their digital agenda forward at a steady pace, but the Covid-19 pandemic required
+a significant leap in the 
+development of digital products and services, with the goal of
+maintaining and fostering customer engagement.
+However, digitalization has generated many cybersecurity 
+issues and the intensification of cyberattacks all around the world~\cite{HKICS20,PA21}.
+In this scenario, Italy is an interesting case study since it is home
+of mostly small to medium enterprises (SMEs), active in different sectors,
+and can consequently highlight cybersecurity aspects that are different from those
+faced by larger companies of greater influence. This paper presents the status of the
+Italian scenario, to understand how cyberattacks and cybersecurity have fared
+during the ongoing COVID-19 pandemic.
+Section~\ref{sec:DigitalAttacksObservatory} presents the \textit{Digital Attacks Observatory},
+a survey that we have used to collect relevant information about cybersecurity and 
+Italian organizations.
+Sections~\ref{sec:DataDiscussion} and~\ref{sec:TypeAttacks} report
+a detailed discussion of the collected data and observed trends.
+Finally, Section~\ref{sec:Conclusion} concludes by underlying the importance of
+a cybersecurity culture focused on the Italian reality.
 
@@
 The \textit{Digital Attacks Observatory} (OAD)~\cite{oadweb} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
-bodies operating and in addition drawn up with the precious collaboration of the Italian Postal and Telecommunications Police. The OAD survey does not provide 
-for a predefined set of respondents, but it allows potential interested parties full and free access to an online questionnaire, in a totally anonymous manner. 
-Thanks to the number of responses collected and their balanced distribution between companies and public bodies of various sizes and belonging to various product 
-sectors, we can provide a specific picture of the cyber attacks in Italy.  The last OAD 2020  reaches the twelfth year of consecutive surveys on cyberattacks in Italy.
+bodies. It is implemented in collaboration with the Italian postal and telecommunications police.
+The last OAD 2020 was the twelfth year of consecutive surveys on cyberattacks in Italy.
+The OAD survey is not directed at a predefined set of respondents, but allows potential interested
+parties to have full and free access to an online questionnaire, in a totally anonymous way. 
+Thanks to the number of responses collected and their balanced distribution between companies and public bodies of different size and belonging to different economic
+sectors, we can provide a specific picture of the cyberattacks in Italy.
 
-\subsection{Questionaries}
+\subsection{The Questionnaires}
 
-The OAD questionaries refer for about half of the question to cyber attacks typologies and they main characteristics, if they have been relived. 
-The other half of the questions refers to the technical and organizational security measures implemented on the IT systems of the respondents. 
-OAD therefore allows to contextualize the cyber attacks with the cyber measures in production in each IT system considered. 
-In fact, OAD provided a macro qualitative evaluation of the implemented measures (within the specific context of the company), in order to improve the motivation to 
-complete the questionaries to all the potential anonymous respondents. 
-In particular the security measures considered in the 2020 questionnaire are subdivided in technical, organizational, management and governance ones as follow:
+About half of the OAD questionnaires refer to the occurred
+cyberattack typologies and their main characteristics.
+The other half refers to the technical and organizational security measures implemented
+on the IT systems of the respondents. 
+Therefeore, OAD allows one to match the cyberattacks to the cyber-measures in each production
+IT system considered. 
+In fact, OAD provides a qualitative macro-evaluation of the implemented measures
+(within the specific context of the company), in order to motivate the anonymous respondents
+to complete the questionnaires.
+In particular, the security measures considered in the 2020 questionnaire were subdivided
+in \emph{technical}, \emph{organizational}, \emph{managerial} and \emph{governamental}, as follow:
+
 \begin{enumerate}
@@
 	\begin{itemize}
-		\item Overall architecture of digital security measures, integrated with the entire IT system architecture
+		\item Overall architecture of the digital security measures, integrated with the entire IT system architecture
 		\item Physical countermeasures
 		\item Identification, Authentication, and Authorization (IAA) measures
-		\item Cyber security measures at local network and geographic level
-		\item Measures for logical protection of each single Information and Communications Technology (ICT) system
+		\item Cybersecurity measures at local and geographical network level
+		\item Measures for logical protection of each single Information and Communication Technology (ICT) system
 		\item Application and software protection
@@
 	\end{itemize}
-	\item \emph{Management and governance measures}
+	\item \emph{Managerial and governamental measures}
 	\begin{itemize}
 		\item Digital security control, monitoring and management systems
-		\item Disaster Recovery (DR) plan
+		\item Disaster Recovery (DR) plan.
 	\end{itemize}
@@
 
-\subsection{Overview 2019-(1Q)2020}
+\subsection{Overview from 2019 to 1Q-2020}
 
-OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
-(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{oad20} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
-of 2020, when in Italy the pandemic exploded. The data comparison of 2019 and the first quarter 2020, provided by the same respondents pool, has a statistical relevance and provides a 
-clear indication of the Covid-19 pandemic on the intentional cyber attacks in Italy. This pandemic has been the trigger for a wide range of cyber attacks, mainly 
-caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
-derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
-even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool is well balanced for the size of the organizations, in terms
-of employees, between those below 250 and the largest ones. It must be taken into account, as reported by ISTAT~\cite{istat21}, that in Italy 99.91\%
-of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
-tiny organizations, which are thee vast majority in Italy, and which are not normally considered in other cyber security surveys: 57.5\% of the OAD 2020 respondents
-belong to structures with less than 250 employees, and of these 22.1\% under 10.
+The OAD 2020 survey
+\footnote{The report is written in Italian,
+  only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
+  (147 A4 pages) and 9 Attachments (39 A4 pages).}~\cite{oad20}
+took place during the
+Covid-19 pandemic and covered the entire 2019 and the first quarter 
+of 2020, when the pandemic exploded in Italy. The comparison of data from 2019 to that
+from the first quarter of 2020, provided by the same set of respondents,
+has statistical relevance and provides a 
+clear indication of the Covid-19 pandemic on the intentional cyberattacks in Italy.
+This pandemic has triggered a wide range of cyberattacks, mainly 
+caused by the sudden, and partly unexpected, passage to working from home
+and to a strong use of every type of IT service on the Internet, as a consequence
+of the mobility lockdowns imposed by the Italian authorities. The resulting respondents pool
+covers all the production sectors, including public administration,
+even if the majority of the respondents companies belong to the ICT sector (30.8\%).
+The 2020 pool is well balanced for the size of the organizations, in terms
+of employees, between those below 250 and the largest ones. It must be taken into account,
+as reported by ISTAT~\cite{istat21}, that 99.91\%
+of enterprises in Italy are SMEs, with up to 250 employees, and 95\% have fewer than ten employees.
+The OAD 2020 survey is therefore able to consider small and 
+tiny organizations, that are the vast majority in Italy, and that usually
+are not considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
+belong to structures with less than 250 employees and, of these, 22.1\% have fewer than ten
+employees.
 
-\section{Data discussion}\label{sec:DataDiscussion}
+\section{Discussion}\label{sec:DataDiscussion}
 
@@
 		\includegraphics[width=1\textwidth]{pictures/fig1.png}
-		\caption{Extract of OAD 2020 \cite{oad20} on macro evaluation of security level in the information systems in Italy.
-The chart shows that more than $ 2/3 $ of the respondents has a good level of cybersecurity.}
+		\caption{Extract of the OAD 2020~\cite{oad20} on the macro-evaluation
+                  of the security level in the IT systems in Italy.
+                  The chart shows that more than two thirds of the respondents
+                  judge the level of cybersecurity of their IT systems as good.}
 		\label{fig:1}
@@
 
-The data collected in the OAD 2020 showed at a high level for the most part a good perception by the respondents of their IT security (Fig.~\ref{fig:1}).
-This information can be explainable from the fact that:
+The data collected in the OAD 2020 show, in general, that the respondents have, for the most part,
+a good perception of the cybersecurity of their IT systems (Fig.~\ref{fig:1}).
+This information can be explained from the fact that:
 
 \begin{itemize}
-\item About 30\% of respondents are in the ICT sector, therefore the IT Systems already known the cybersecurity culture and can properly react to the cyber attacks, preventing 
-attacks and minimizing risks.
+\item About 30\% of respondents are in the ICT sector, therefore their IT systems
+  are managed with some cybersecurity literacy and can properly react to cyberattacks, preventing 
+  them or minimizing their effects.
 
-\item In general, cybercriminals do not waste time in attacks targeting small businesses, to obtain only a marginal profit, 
-with the risk of being caught by the police.  As confirmed by trends of Fig.~\ref{fig:3}, the attackers try to hit the big structures,
- because with these it is more probable the realization of gains, even if illegal.
+\item In general, cybercriminals do not waste time and resources to attack small businesses,
+  where they can only obtain a marginal profit and still
+  take the risk of being caught by the police.  As confirmed by the trends shown
+  in Fig.~\ref{fig:3}, attackers target the larger companies most,
+  because the illegal gain is more interesting in that case.
 \end{itemize}
@@
   \centering
+  \includegraphics[width=1\textwidth]{pictures/fig3.png}
+  \caption{Comparison between cyberattacks occurred in 2019 and in the first quarter of 2020,
+    wrt.\ to the organization dimension~\cite{oad20}. The dotted line links
+    the unreported attacks and shows that they decrease while moving from tiny to big organizations,
+    in terms of employees.}
+  \label{fig:3}
+\end{figure}
+
+\begin{figure}
+  \centering
   \includegraphics[width=1\textwidth]{pictures/fig2.png}
-		\caption{Trend comparison of the relived attacks from the beginning of the OAD survey in 2008~\cite{oad20}. This comparison is only valid as a trend and not from a statistical point 
-of view, due to the fact that the respondents pools year by year are different.
-The dotted red line shows a wave shape: after a relative pick of attacks, the companies react hardening their cyber security, and the following year, normally,
-the relative pick in \% is lower that the previous one.}
+  \caption{Trend of the reported attacks from the beginning of the OAD survey
+    in 2008~\cite{oad20}. This comparison is only valid as a trend and not as a statistical
+    measure, since the respondents are different from year to year. The dotted red line shows
+    successive waves: after a relative pick of attacks, companies react and strengthen
+    their cybersecurity measures, so that, the following year, the trend dicreases.}
   \label{fig:2}
@@
 
-Analyzing the trends about cyber attaks since 2007 to 2015, the chart in Fig.~\ref{fig:2} reveals in average that about 40\% of the respondents that relived cyber attacks 
-(first red line). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line). 
-In 2018 OAD reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that percentage of the not relived. After the biggest pick 
-in 2018, 2019 has a lower percentage, and also for the first quarter 2020. Only with the 2021 survey we will discover the percentages for the whole 2020. 
-This view over several years of the interviewees does not show significant changes especially compared to the pandemic period but only a slight increase in cyber attacks 
-in recent years not necessarily attributable to the pandemic crisis.
+The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattaks.
+The figure shows that, from 2007 to 2015, in average, about 40\% of the respondents
+reported cyberattacks (leftmost full red line). Since 2015, the trend increased to 45\%
+of the respondents (rightmost full red line). 
+In 2018, OAD reached the pick of reported attacks and, for the first time, the percent of
+reported attacks became larger than that percent of companies that did not report any attack.
+After the pick 
+in 2018, year 2019 has a lower figure and the same occurs in the first quarter 2020.
+The trend for full 2020 will be discovered with the currently ongoing OAD survey.
+This view, over several years, does not show significant changes, not even inside
+the pandemic period, but shows only a slight increase in cyberattacks 
+in the recent years, not necessarily attributable to the pandemic crisis.
 
-\begin{figure}
-	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig3.png}
-		\caption{Comparison between cyber attacks relived in 2019 and first quarter of 2020 for the organization dimension~\cite{oad20}. The black dotted line in the figure shows 
-the “not” relieved attacks, and points out how it decrease from the tiny to the big organizations, in terms of employees. }
-		\label{fig:3}
-\end{figure}
+As already highlighted above, the comparison between the whole 2019 and the first quarter 2020
+in Fig.~\ref{fig:3} maintains the same logical trend. The only meaningful difference
+is in the number of attacks reported by the smallest
+organizations (up to ten employees): they were 0\% in 2019 and grwo to
+22,2\% at the beginning of 2020 (see the red arrow in Fig.~\ref{fig:3}).
+Such a strong increase follows from the extensive use of e-banking and 
+e-commerce services, with the related cyberattacks.
 
-As already highlighted above, the comparison between the whole 2019 and the first quarter 2020 in Fig.~\ref{fig:3} maintains the same logical trend.
-The only meaningful difference is for the frequent attacks relieved by the very
-small organizations, up to 10 employees: 0\% in 2019 grows up to 22,2\% (see in Fig.~\ref{fig:3} the red line). Such a strong increase derives by the large utilization of e-baking and 
-e-commerce services, with the related cyber attacks.
-
-The impact of pandemic on the cyber security in Italy is also highlighted by the data provided by the Postal and Communications Police that shows the situation for the
-critical infrastructures (Table.\ref{tab:table1}), for the financial cyber crimes (Table.\ref{tab:table2}) and for the cyber terrorism (Table.\ref{tab:table3}). 
-Fraudulent transactions in 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the 1st quarter of 
-2020 fraudulent transactions for \euro 20.2 million were blocked, practically reaching in 4 months what was blocked in the 12 months of the previous year, and \euro 8 million 
-recovered; an indicator of the growth in attacks on financial transactions and services due to the very large use of these online services caused by the mobility blocks imposed 
-by the Covid-19 pandemic.
+The impact of the pandemic on the cybersecurity in Italy is also highlighted by the data
+provided by the postal and communication police, that shows the situation for the
+critical infrastructures (Table~\ref{tab:table1}), for the financial cyber-crimes
+(Table~\ref{tab:table2}) and for the cyber-terrorism (Table~\ref{tab:table3}). 
+Fraudulent transactions in 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the first quarter of 
+2020, fraudulent transactions for \euro 20.2 million were blocked,
+practically reaching in four months the amount
+blocked in the full twelve months of 2019, and \euro 8 million have been
+recovered. This is an indication of the increase in attacks on financial transactions
+and services due to the very large use of these online services, caused by the mobility reduction
+measures consequent of the Covid-19 pandemic.
 
@@
   \begin{center}
-  \resizebox{1.15\textwidth}{!}{
+  \resizebox{1\textwidth}{!}{
 	\begin{tabular}{|r|r|r|r|r|r|}									
 		\hline
-		Critical structure protection	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018 &	1 Jan - 31 Dec 2017 &	1 Jan - 31 Dec 2016\\
+		\textbf{Critical structure protection}	&	1 Jan - 30 Apr	&	1 Jan - 31 Dec	&	1 Jan - 31 Dec &	1 Jan - 31 Dec &	1 Jan - 31 Dec\\
+                	&	2020	&	2019	&	2018 &	2017 &	2016\\
 		\hline									
@@
 		\hline	
-		Initiated investigations on detected attacks	&	12,05\%	&	13,12\%	&	16,12\%	&	6,98\%	& 8,29\% \\
+		Initiated investigations	&	12,05\%	&	13,12\%	&	16,12\%	&	6,98\%	& 8,29\% \\
 		\hline	
-		People arrested out of	reported &	0\%	&	2,5\%	&	7,14\%	&	0,23\%	&	0,24\%\\						
+		People arrested out ofreported &	0\%	&	2,5\%	&	7,14\%	&	0,23\%	&	0,24\%\\						
 		\hline	
@@
 
-	\caption{Data of critical structure protection collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\caption{Data of critical structure protection collected by Italian postal and communication police~\cite{oad20}.}									
 	\label{tab:table1}									
@@
   \begin{center}
-  \resizebox{1.15\textwidth}{!}{
+  \resizebox{1\textwidth}{!}{
 	\begin{tabular}{|r|r|r|r|r|r|}									
 		\hline									
-		Financial Cyber Crime	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018 &	1 Jan - 31 Dec 2017 &	1 Jan - 31 Dec 2016\\
+		\textbf{Financial Cyber Crime}	&	1 Jan - 30 Apr	&	1 Jan - 31 Dec	&	1 Jan - 31 Dec &	1 Jan - 31 Dec &	1 Jan - 31 Dec\\
+                	&	2020	&	2019	&	2018 &	2017 &	2016\\
 		\hline									
 		Blocked fraudulent transactions	&	\textbf{\euro 20.200.000,00}	&	\textbf{\euro 21.333.990,00}	&	\euro 38.400.000,00	& \euro 20.839.576,00 & \euro 16.050.812,50 	\\
-		Recovered amounts 	&	\euro 8.700.000,00	&	\euro 18.000.000,00	&	\euro 9.000.000,00	&	\euro 862.000,00 & n.d.	\\
-		People arrested	&	n.d.	&	n.d.	&	n.d.	&	25	& 25\\
-		People reported	&	n.d.	&	n.d.	&	n.d.	&	2.851 & 3.772	\\
+		Recovered amounts 	&	\euro 8.700.000,00	&	\euro 18.000.000,00	&	\euro 9.000.000,00	&	\euro 862.000,00 & n.a.	\\
+		People arrested	&	n.a.	&	n.a.	&	n.a.	&	25	& 25\\
+		People reported	&	n.a.	&	n.a.	&	n.a.	&	2.851 & 3.772	\\
 		\hline	
@@
     \end{center}
-	\caption{Data of critical structure protection collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\caption{Data of critical structure protection collected by Italian postal and communication police~\cite{oad20}.}									
 	\label{tab:table2}									
@@
 		\hline									
-		Cyber terrorism	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018\\
+		\textbf{Cyber-terrorism}	&	1 Jan - 30 Apr 2020	&	1 Jan - 31 Dec 2019	&	1 Jan - 31 Dec 2018\\
 		\hline									
 		Monitored web space	&	\euro 11.962	&	\euro 36.377	&	\euro 36.000\\
-		Removed contents 	&	n.d.	&	n.d.	&	250\\
+		Removed contents 	&	n.a.\	&	n.a.\	&	250\\
 		\hline	
@@
 
-\section{Cyber attack trends}\label{sec:TypeAttacks}
+\section{Nature of Cyberattacks}\label{sec:TypeAttacks}
 
-In the pandemic period between 2019 and 1st quarter of 2020, the attacks on IAA that aimed at access control systems are most widespread with 34\% in 2019 and 28.2\% (Fig.~\ref{fig:5}).
-As diffusion among the respondents, the other 14 types of attacks follow, decreasing by a few percentage points among them, whose characteristics and impacts are detailed in the related 
-paragraphs of the OAD 2020 report. Regarding attack vectors and techniques, all the seven attacks families considered in the 2020 questionnaire are widely used in the various 
-types of attacks, some times even simultaneously., as shown in Fig.~\ref{fig:6}. The most widespread one,
-as an average on the various attacks detected by the respondents’ pool, is the use of toolkits (rootkits, meta-exploits, etc.) for the identification and exploitation of vulnerabilities
-on the target system of the attack, with a 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.) with 34.6\% 
-and the use of malicious code and scripts with 34.3\%. The other considered technolog decrease with a few percentage points of difference between them. The impacts of the most 
-critical attacks are analyzed for each attack type, as well as their possible motivations and the recovery times required by the most critical ones.
-In the 2020 OAD survey, all these attacks’ characteristics vary for each type of attack, and the emerged results are described in the specific paragraphs dedicated to each attack 
-type. At overall it emerges that:
+In the pandemic period, between 2019 and the first quarter of 2020, the nature of the attacks and their vectors were quite diversified.
+The attacks on IAA that aimed at access control systems are widespread, since they amount to 28.2\% of the total attacks
+of 2019 and to 34\% of the total attacks in the first quarter of 2020 (Fig.~\ref{fig:5}).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1\textwidth]{pictures/fig5.png}
+  \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.}
+  \label{fig:5}
+\end{figure}
+
+The other 14 types of attacks follow, decreasing by a few percent points between them, whose characteristics and impacts are detailed in the related 
+paragraphs of the OAD 2020 report. Regarding attack vectors and techniques, all the seven attack families considered in the 2020 questionnaire are widely used in the various 
+types of attacks, sometime even simultaneously, as shown in Fig.~\ref{fig:6}. The trend in Fig.~\ref{fig:6} is also in line with most European reports, such as
+that of the European Network and Information Security Agency (ENISA)~\cite{enisa20}. 
+The most widespread one, in the average of the various attacks detected by the respondents, is the use of toolkits (rootkits, meta-exploits, etc.)
+for the identification and exploitation of vulnerabilities
+on the target system, with 38.8\%, followed by the well-known malicious and unauthorized collection of information (social engineering, phishing, etc.), with 34.6\%,
+and the use of malicious code and scripts, with 34.3\%. The other considered techniques decrease, with a few percent points between them. The impacts of the most 
+critical attacks are analyzed, for each attack type, as well as their possible motivations and the recovery time required by the most critical ones.
+In the 2020 OAD survey, all these attack characteristics vary for each type of attack, and the emerged results are described in the specific paragraphs dedicated to each attack 
+type. In the overall, it emerges that:
 \begin{itemize}
-\item the impacts declared by the respondents are balanced between those irrelevant and/or easily resolvable with recovery in short time and at limited costs, and those very 
-critical, which require expensive and long recovery actions, and that, in some cases, can cause business and customer loses; these two very different impact cases depend mainly 
+\item the impacts declared by the respondents are balanced between those irrelevant and/or easily resolvable with a quick recovery and at limited costs, and those very 
+critical, that require expensive and long recovery actions and that, in some cases, can cause business and customer losses. These two very different impact cases depend mainly 
 on the security measures in place;
-
-\item the motivations of the cyber attacks are mainly of an economic nature, therefore done for fraud and blackmail: the widespread diffusion of ransomware in Italy is a clear 
+\item the motivations of the cyberattacks are mainly economic, such as fraud and blackmailing: the widespread diffusion of ransomware in Italy is a clear 
 confirmation of these motivations.
@@
   \centering
-		\includegraphics[width=1\textwidth]{pictures/fig5.png}
-		\caption{Diffusion of digital attacks on respondents' computer systems based on the type of attack between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
-		\label{fig:5}
-\end{figure}
-\begin{figure}
-	\centering
   \includegraphics[width=0.90\textwidth]{pictures/fig6.png}
-		\caption{Average diffusion (percentage) of the attack techniques detected in the most serious attacks between 2019 - 1st quoter 2020 (multiple answers) \cite{oad20}}
+  \caption{Average diffusion (in percent) of the attack techniques detected in the most serious attacks between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.}
   \label{fig:6}
@@
 
-As pointed out in Fig.~\ref{fig:1}, the Informatics Systems considered in the 2020 survey and their cybersecurity  are in the medium-high range and the collected information show a relevant 
-improvement in both technical and organizational measures, in comparison to the previous surveys. Few of the considered Italian Informatics Systems are based on Data Centers in 
-Italy, most of the respondents’ companies have medium-small Informatic Systems partly on premise and partly outsourced, with an increasing use of cloud services.
-The high number of attacks detected in 2018, and the privacy obligations, related to General Data Protection Regulation (GDPR), have certainly contributed to strengthening cyber security measures, and a further improvement
-comes from the growing use of cloud services, where usually there are high standard security measures. Organizational measures for cybersecurity, historically lacking and neglected
-in Italy, have improved in terms of defining roles and separation of duties, of organizational policies and procedures, and of incidents management. These measures often are lacking
-in small organizations, and in general the cybersecurity awareness and competence are low, and mainly at the top level of the public and private organizations. In Italy, there is 
-still a long way to go in terms of cybersecurity continual training and awareness. A formal and bureaucratic approach for the organizational procedures is still prevalent, which 
-once defined, often do not find a possible concrete applicability, periodic updates and operational tests: an emblematic example is represented by the Disaster Recovery plans for 
-which, several times, the required alternative ICT resources to be used are not forecasted and allocated and there also not provided periodic exercises and simulations of the 
-possible disasters. Cybersecurity management tools still have limited diffusion among the 2020 OAD respondents, in particular the most advanced ones based on artificial intelligence
-techniques.
-Although the use of IoT \cite{JK21}, industrial automation, robotics systems \cite{SGLMDXHKCZT21}, and systems based on blockchain technologies \cite{RMFF21} are expanding and can allow to combat the Covid-19 pandemic, in
-the OAD 2020, we can find only a low number of respondents involved, which also derives from the limited percentage of their product sectors that should be the more interested in these systems: the manufacturing sectors, the logistics, the research and
-development centers and labs, the Local Public Administrations for territorial control, and so on. For these topics, the data that emerged from the survey are considered only as 
-specific cases that contribute to the general values on the cyberattacks, but which cannot be still considered of reference at the Italian level.
+As pointed out in Fig.~\ref{fig:1}, the IT systems considered in the 2020 survey and their cybersecurity are in the medium to high range and the collected information shows a relevant 
+improvement in both technical and organizational measures, in comparison to the previous surveys. Few of the considered Italian IT systems are based on data centers in 
+Italy: most of the respondents' companies have medium to small IT systems partly on premise and partly outsourced, with an increasing use of cloud services.
+The high number of attacks detected in 2018, and the privacy obligations related to the European
+General Data Protection Regulation (GDPR), have certainly contributed to strengthening cybersecurity measures, and a further improvement 
+comes from the growing use of cloud services, where high standard security measures are in place, usually. Organizational measures for cybersecurity, historically missing and neglected
+in Italy, have improved in terms of defition of roles and separation of duties, of organizational policies and procedures, and of incident management. These measures often are missing
+in small organizations, and in general the cybersecurity awareness and competence are low, and mainly concentrated at the top level of the public and private organizations. In Italy, there is 
+still a long way to go in terms of cybersecurity continued training and awareness. A formal and bureaucratic approach to the organizational procedures is still prevalent. Once defined,
+such procedures often do not find concrete application and misses periodic updates and operational tests. A significant example is the disaster recovery plans, for 
+which, typically, the required alternative IT resources are neither forecast nor allocated. Periodic exercises and simulations of the 
+possible disasters are not implemented. Cybersecurity management tools have limited diffusion yet, among the 2020 OAD respondents. This is particularly true for
+the most advanced ones, based on artificial intelligence.
+Although the use of IoT~\cite{JK21}, industrial automation, robotics systems~\cite{SGLMDXHKCZT21} and systems based on blockchain technology~\cite{RMFF21}
+are expanding and can contribute to fight the Covid-19 pandemic.
+In the OAD 2020 report, one finds only a few respondents involved in such technologies, which also derives from the limited portion
+of their economic sector, interested in such systems: the manufacturing sectors, the logistics, the research and 
+development centers and labs, the local public administration for territorial control, and so on.
+%For these topics, the data that emerged from the survey are considered only as 
+%specific cases that contribute to the general values on the cyberattacks, but which cannot be still considered of reference at the Italian level.
 
@@
 
-The OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of an high complexity and sophistication, with a slightly increasing 
-spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 
-degrees related to the Covid-19 pandemic, in the first four months of 2020 the general spread of attacks is at similar values to those of 2019: to be verified with the next OAD 2021
-if there will be changes to this trend.
-The Italian reality, made up of a very large number of small and tiny organizations, does not make our country one of the most attractive for cybercriminals, but cyber warfare
-and massive attacks represent a growing and serious risk, as has already happened in part with the widespread of ransomware on computer systems whit a lack of or  with low 
-cybersecurity  basic measures. OAD 2020 notes a clear improvement and strengthening of digital security measures, both technical and organizational, even if the most modern 
-prevention, protection and management systems that use artificial intelligence techniques are still embryonic  in the respondents' basin. The defense measures and techniques 
-in use chase the increasingly sophisticated and smart evolution of attacks, but almost always late. The high density of vulnerabilities 
-requires different approaches and new logics, with the aim of making all ICT systems interconnectable to the Internet intrinsically safe, by default and by design. But we are 
-still a long way from this goal, and in order to decisively improve the concrete fight against continuous attacks and cybercrime, it is currently necessary to increase 
-cybersecurity’ awareness and skills at all levels, an effective collaboration between the police at world level, and primarily a real and large usage of the professional ethics
- both of those involved (supply side) and both of those who decide (demand side) on cybersecurity.
+The OAD 2020 survey highlights the presence of cyberattacks of various types in Italy, all technically of an high complexity and sophistication, with a slight increase 
+compared to the trend that emerged in the twelve years of the previous OAD surveys, but still below the pick of 2018. Despite the proliferation of cyberattacks
+related to the Covid-19 pandemic, in the first four months of 2020 the widespread of attacks is at a level similar to that of 2019. This will be better verified
+with the next OAD 2021. The Italian reality, made up of a very large number of small and tiny organizations, does not make our country very attractive to cybercriminals, but cyber-warfare
+and massive attacks represent a growing and serious risk, as has already occurred with the widespread of ransomware on computer systems with no or low
+measures of cybersecurity. The OAD 2020 shows a clear improvement and strengthening of digital security measures, both technical and organizational, even if the most modern 
+prevention, protection and management systems that use artificial intelligence techniques are still in their infancy among the respondents. The defense measures and techniques 
+in use fight the increasingly sophisticated and smart evolution of attacks, but often too late. The high density of vulnerabilities 
+requires different approaches, aiming at making all IT systems interconnected to the Internet and safe, by default and by design. But we are 
+still a long way from this goal. In order to improve the actual fight against continuous attacks and cybercrime, it is currently necessary to increase 
+awareness on cybersecurity and skills at all levels. It is also paramount to support an effective collaboration between police bodies at the international level.
 

### 2021-03-08 article.tex
Added email
Updated pictures
Added numbers of responses

@@
 \and
-  Digital Attacks Observatory (OAD) Team, Italy
+Digital Attacks Observatory (OAD) Team, Italy\\
+\email{m.bozzetti@aipsi.org}
 \and
@@
 The \textit{Digital Attacks Observatory} (OAD)~\cite{oadweb} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
-bodies. It is implemented in collaboration with the Italian postal and telecommunications police.
+bodies. It is implemented in collaboration with the Italian postal and telecommunication police.
 The last OAD 2020 was the twelfth year of consecutive surveys on cyberattacks in Italy.
@@
 parties to have full and free access to an online questionnaire, in a totally anonymous way.
+The survey has collected responses from 310 enterprises in 2020, in comparison
+to 296 and 2019 and 269 in 2018.
 Thanks to the number of responses collected and their balanced distribution between companies and public bodies of different size and belonging to different economic
-sectors, we can provide a specific picture of the cyberattacks in Italy.
+sectors, we can provide now a specific picture of the cyberattacks in Italy.
 
@@
   \centering
-  \includegraphics[width=1\textwidth]{pictures/fig3.png}
+  \includegraphics[width=1\textwidth]{pictures/fig2.png}
+  \caption{Trend of the reported attacks from the beginning of the OAD survey
+    in 2008~\cite{oad20}. This comparison is only valid as a trend and not as a statistical
+    measure, since the respondents are different from year to year. The dotted red line shows
+    successive waves: after a relative peak of attacks, companies react and strengthen
+    their cybersecurity measures, so that, the following year, the trend dicreases.}
+  \label{fig:2}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.52]{pictures/fig3.png}
+	\end{center}
   \caption{Comparison between cyberattacks occurred in 2019 and in the first quarter of 2020,
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=1\textwidth]{pictures/fig2.png}
-  \caption{Trend of the reported attacks from the beginning of the OAD survey
-    in 2008~\cite{oad20}. This comparison is only valid as a trend and not as a statistical
-    measure, since the respondents are different from year to year. The dotted red line shows
-    successive waves: after a relative pick of attacks, companies react and strengthen
-    their cybersecurity measures, so that, the following year, the trend dicreases.}
-  \label{fig:2}
-\end{figure}
-
 The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattaks.
@@
 of the respondents (rightmost full red line). 
-In 2018, OAD reached the pick of reported attacks and, for the first time, the percent of
+In 2018, OAD reached the peak of reported attacks and, for the first time, the percent of
 reported attacks became larger than that percent of companies that did not report any attack.
-After the pick 
+After the peak 
 in 2018, year 2019 has a lower figure and the same occurs in the first quarter 2020.
@@
 		\hline	
-		People arrested out ofreported &	0\%	&	2,5\%	&	7,14\%	&	0,23\%	&	0,24\%\\						
+		People arrested out of reported &	0\%	&	2,5\%	&	7,14\%	&	0,23\%	&	0,24\%\\						
 		\hline	
@@
         \end{center}
-	\caption{Data of cyber terrorism collected by Italian Postal and Communications Police~\cite{oad20}}									
+	\caption{Data of cyber terrorism collected by Italian postal and communication Police~\cite{oad20}.}									
 	\label{tab:table3}									
@@
 The OAD 2020 survey highlights the presence of cyberattacks of various types in Italy, all technically of an high complexity and sophistication, with a slight increase 
-compared to the trend that emerged in the twelve years of the previous OAD surveys, but still below the pick of 2018. Despite the proliferation of cyberattacks
+compared to the trend that emerged in the twelve years of the previous OAD surveys, but still below the peak of 2018. Despite the proliferation of cyberattacks
 related to the Covid-19 pandemic, in the first four months of 2020 the widespread of attacks is at a level similar to that of 2019. This will be better verified

### 2021-03-09 article.tex
Minor changes

@@
 the pandemic period, but shows only a slight increase in cyberattacks 
-in the recent years, not necessarily attributable to the pandemic crisis.
+in the recent years, not necessarily attributable to the pandemic crisis. One could wonder
+why the percentages in Fig.~\ref{fig:2} are lower than what the reader would have expected, 
+given the numerous cyber attacks occured in Italy. As already discussed at the beginning of 
+this section, the main reason can be attributed to the sizes of italian companies.
 
-As already highlighted above, the comparison between the whole 2019 and the first quarter 2020
+The comparison between the whole 2019 and the first quarter 2020
 in Fig.~\ref{fig:3} maintains the same logical trend. The only meaningful difference
 is in the number of attacks reported by the smallest
-organizations (up to ten employees): they were 0\% in 2019 and grwo to
+organizations (up to ten employees): they were 0\% in 2019 and grow up to
 22,2\% at the beginning of 2020 (see the red arrow in Fig.~\ref{fig:3}).
@@
 		People reported	&	0	&	117	&	14	&	1.316 &	1.226	\\
-		Perquisitions	&	n.d.	&	n.d.	&	n.d.	&	73	& 58 \\
+		Perquisitions	&	n.a.	&	n.a.	&	n.a.	&	73	& 58 \\
 		\hline	
@@
   \includegraphics[width=1\textwidth]{pictures/fig5.png}
-  \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.}
+  \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.
+(*) Non all the IT systems of the respondents utilize these types of solutions/systems}
   \label{fig:5}

### 2021-03-09 article.tex
Small final edits

@@
 After the peak 
-in 2018, year 2019 has a lower figure and the same occurs in the first quarter 2020.
+in 2018, year 2019 has a lower figure and the same occurs in the first quarter of 2020.
 The trend for full 2020 will be discovered with the currently ongoing OAD survey.
@@
 given the numerous cyber attacks occured in Italy. As already discussed at the beginning of 
-this section, the main reason can be attributed to the sizes of italian companies.
+this section, the main reason seems the size of the Italian companies.
 
-The comparison between the whole 2019 and the first quarter 2020
+The comparison between the whole 2019 and the first quarter of 2020
 in Fig.~\ref{fig:3} maintains the same logical trend. The only meaningful difference
@@
   \caption{Widespread of the digital attacks on the respondents' computer systems, classified wrt.\ the type of attack, between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.
-(*) Non all the IT systems of the respondents utilize these types of solutions/systems}
+Where marked with (*), we note that not all the IT systems of the respondents use these types of solutions/systems.}
   \label{fig:5}

### 2021-03-09 article.tex
Minor format change
Fixed police reference

@@
 The \textit{Digital Attacks Observatory} (OAD)~\cite{oadweb} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
-bodies. It is implemented in collaboration with the Italian postal and telecommunication police.
+bodies. It is implemented in collaboration with the Italian Postal and Telecommunication Police.
 The last OAD 2020 was the twelfth year of consecutive surveys on cyberattacks in Italy.
@@
 	\centering
-		\includegraphics[width=1\textwidth]{pictures/fig1.png}
+		\includegraphics[width=0.9\textwidth]{pictures/fig1.png}
 		\caption{Extract of the OAD 2020~\cite{oad20} on the macro-evaluation
@@
 The impact of the pandemic on the cybersecurity in Italy is also highlighted by the data
-provided by the postal and communication police, that shows the situation for the
+provided by the Italian Postal and Telecommunication Police, that shows the situation for the
 critical infrastructures (Table~\ref{tab:table1}), for the financial cyber-crimes
@@
 
-	\caption{Data of critical structure protection collected by Italian postal and communication police~\cite{oad20}.}									
+	\caption{Data of critical structure protection collected by Italian Postal and Telecommunication Police~\cite{oad20}.}									
 	\label{tab:table1}									
@@
     \end{center}
-	\caption{Data of critical structure protection collected by Italian postal and communication police~\cite{oad20}.}									
+	\caption{Data of critical structure protection collected by Italian Postal and Telecommunication Police~\cite{oad20}.}									
 	\label{tab:table2}									
@@
         \end{center}
-	\caption{Data of cyber terrorism collected by Italian postal and communication Police~\cite{oad20}.}									
+	\caption{Data of cyber terrorism collected by Italian Postal and Telecommunication Police~\cite{oad20}.}									
 	\label{tab:table3}									

### 2021-03-09 article.tex
Fixed typos

@@
   industrial scenario is mostly populated by small and medium enterprises,
-  that have been obliged to a quick reconvertion of their IT systems and typically
+  that have been obliged to a quick reconversion of their IT systems and typically
   lack the necessary cybersecurity culture.
@@
 on the IT systems of the respondents. 
-Therefeore, OAD allows one to match the cyberattacks to the cyber-measures in each production
+Therefore, OAD allows one to match the cyberattacks to the cyber-measures in each production
 IT system considered. 
@@
 	\end{itemize}
-	\item \emph{Managerial and governamental measures}
+	\item \emph{Managerial and governmental measures}
 	\begin{itemize}
@@
     successive waves: after a relative peak of attacks, companies react and strengthen
-    their cybersecurity measures, so that, the following year, the trend dicreases.}
+    their cybersecurity measures, so that, the following year, the trend decreases.}
   \label{fig:2}
@@
 
-The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattaks.
+The chart in Fig.~\ref{fig:2} analyzes the trend of cyberattacks.
 The figure shows that, from 2007 to 2015, in average, about 40\% of the respondents
@@
 why the percentages in Fig.~\ref{fig:2} are lower than what the reader would have expected, 
-given the numerous cyber attacks occured in Italy. As already discussed at the beginning of 
+given the numerous cyber attacks occurred in Italy. As already discussed at the beginning of 
 this section, the main reason seems the size of the Italian companies.
@@
 comes from the growing use of cloud services, where high standard security measures are in place, usually. Organizational measures for cybersecurity, historically missing and neglected
-in Italy, have improved in terms of defition of roles and separation of duties, of organizational policies and procedures, and of incident management. These measures often are missing
+in Italy, have improved in terms of definition of roles and separation of duties, of organizational policies and procedures, and of incident management. These measures often are missing
 in small organizations, and in general the cybersecurity awareness and competence are low, and mainly concentrated at the top level of the public and private organizations. In Italy, there is 

### 2021-03-29 article.tex
Added review about sectors' comments

@@
 The survey has collected responses from 310 enterprises in 2020, in comparison
-to 296 and 2019 and 269 in 2018.
-Thanks to the number of responses collected and their balanced distribution between companies and public bodies of different size and belonging to different economic
-sectors, we can provide now a specific picture of the cyberattacks in Italy.
+to 296 and 2019 and 269 in 2018. Thanks to the number of responses collected and their balanced 
+distribution between companies and public bodies of different size, belonging to different economic
+sectors, and different geographical areas, we can provide now a specific picture
+ of the cyberattacks in Italy.
 
@@
 of the mobility lockdowns imposed by the Italian authorities. The resulting respondents pool
-covers all the production sectors, including public administration,
+covers all the production sectors (Fig.~\ref{fig:9}), including public administration,
 even if the majority of the respondents companies belong to the ICT sector (30.8\%).
+However, many of them are represented by low percentages. this reason, not having sufficiently
+representative samples, we have not considered particular data analysis and correlations by sectors.
 The 2020 pool is well balanced for the size of the organizations, in terms
@@
 employees.
+The distribution of annual respondents' turnover also reflects the turnover of SMEs, as shown in Fig.~\ref{fig:7}.
+Regarding geographical distribution (Fig.~\ref{fig:8}), the responders' companies and organizations are located for the most part only on Italian territory (77.9\%), predominantly in northern Italy (43.3\%).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.90\textwidth]{pictures/fig9.png}
+	\caption{Product breakdown of the respondents' companies and organizations~\cite{oad20}.}
+	\label{fig:9}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.90\textwidth]{pictures/fig7.png}
+	\caption{Last annual available turnover of responders' companies and organizations~\cite{oad20}. 
+		 It should be noted that almost 20\% of the respondents cannot or do not want provide the turnover class to which they fall, stating that the confidentiality policies prevent to indicating a turnover class, even in an investigation completely anonymous.
+	}
+	\label{fig:7}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.90\textwidth]{pictures/fig8.png}
+	\caption{Geographical coverage of the companies/organizations of respondents~\cite{oad20}.}
+	\label{fig:8}
+\end{figure}
 
@@
 development centers and labs, the local public administration for territorial control, and so on.
-%For these topics, the data that emerged from the survey are considered only as 
-%specific cases that contribute to the general values on the cyberattacks, but which cannot be still considered of reference at the Italian level.
 

### 2021-03-31 article.tex
Minor fix
Added info of attacks wtr geographic area

@@
 \begin{enumerate}
-\small
+\tiny
 	\item \emph{Technical measures}
@@
 even if the majority of the respondents companies belong to the ICT sector (30.8\%).
-However, many of them are represented by low percentages. this reason, not having sufficiently
+However, many of them are represented by low percentages. For this reason, not having sufficiently
 representative samples, we have not considered particular data analysis and correlations by sectors.
@@
   take the risk of being caught by the police.  As confirmed by the trends shown
-  in Fig.~\ref{fig:3}, attackers target the larger companies most,
+  in Fig.~\ref{fig:3} and Fig.~\ref{fig:4}, attackers target the larger companies most,
   because the illegal gain is more interesting in that case.
@@
   \centering
-  \includegraphics[width=1\textwidth]{pictures/fig2.png}
+  \includegraphics[width=0.9\textwidth]{pictures/fig2.png}
   \caption{Trend of the reported attacks from the beginning of the OAD survey
@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.46]{pictures/fig4.png}
+	\end{center}
+  \caption{Trend of the reported attacks wrt. to the organizations' turnover in 2019.}
+  \label{fig:4}
+\end{figure}
+
+Regarding the geographical distribution of attacks, Table.~\ref{tab:table4} shows
+that occasional and infrequent attacks are evenly distributed throughout the territory in 2019,
+with systematic attacks focused on northern Italy (where the companies with the highest turnover are present). 
+While, Table.~\ref{tab:table5} shows that already in the first months of 2020 there were several attacks.
+
+\begin{table}[ht]
+  \begin{center}
+  \resizebox{1\textwidth}{!}{
+	\begin{tabular}{|l|r|r|r|}									
+		\hline
+		\textbf{Geographic area} &	Occasional cyber attacks	&	Frequent and repeated cyber attacks & Never detected cyber attacks \\
+		& (few dozen in the year)	& (hundreds or more in the year) & \\
+		\hline									
+		Northern Italy	&	76,3\%	&	18,0\%	&	5,7\% \\
+		Central Italy	&	46,5\%	&	0,0\%	&	53,5\%	\\
+		Southern Italy and islands	&	80,8\%	&	0,0\% &	19,2\% \\
+		Throughout Italy &	100,0\%	&	0,0\% &	0,0\% \\
+		Throughout Italy and Europe &	58,8\%	&	0,0\%	&	41,2\%	\\
+		Throughout Italy and abroad	(also outside Europe) &	100,0\%	&	0,0\%	&	0,0\% \\	
+		\hline	
+	\end{tabular}
+  }
+    \end{center}
+
+	\caption{Distribution of digital attacks (detected by the respondents) wrt. to geographic area in 2019. The percentages of central Italy are 
+	very different from the others because the organizations responding to that geographical area were numerically lower than the rest 
+	of the survey participant (Fig.\ref{fig:8}).}									
+	\label{tab:table4}									
+\end{table}
+
+\begin{table}[ht]
+  \begin{center}
+  \resizebox{1\textwidth}{!}{
+	\begin{tabular}{|l|r|r|r|}									
+		\hline
+		\textbf{Geographic area} &	Occasional cyber attacks	&	Frequent and repeated cyber attacks & Never detected cyber attacks \\
+		& (few dozen)	&  (hundreds or more) & \\
+		\hline									
+		Northern Italy	&	5,8\%	&	6,3\%	&	87,9\% \\
+		Central Italy	&	0,0\%	&	0,0\%	&	100,0\%	\\
+		Southern Italy and islands	&	0,0\%	&	0,0\% &	100,0\% \\
+		Throughout Italy &	30,3\%	&	0,0\%	&	69,7\%	\\
+		Throughout Italy and Europe &	0,0\%	&	0,0\%	&	100,0\%	\\
+		Throughout Italy and abroad (also outside Europe)	&	28,8\%	&	0,0\%	&	71,2\%	\\					
+		\hline		
+	\end{tabular}
+  }
+    \end{center}
+
+	\caption{Distribution of digital attacks (detected by the respondents) wrt. to geographic area in 1Q-2020.}									
+	\label{tab:table5}									
+\end{table}
+
 The impact of the pandemic on the cybersecurity in Italy is also highlighted by the data
@@
 (Table~\ref{tab:table2}) and for the cyber-terrorism (Table~\ref{tab:table3}). 
-Fraudulent transactions in 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the first quarter of 
+Fraudulent transactions all Italyin 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the first quarter of 
 2020, fraudulent transactions for \euro 20.2 million were blocked,
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \begin{center}
@@
   \centering
-  \includegraphics[width=0.90\textwidth]{pictures/fig6.png}
+  \includegraphics[width=0.85\textwidth]{pictures/fig6.png}
   \caption{Average diffusion (in percent) of the attack techniques detected in the most serious attacks between 2019 and the first quarter of 2020 (multiple answers)~\cite{oad20}.}

### 2021-03-31 article.tex
Small edits

@@
 even if the majority of the respondents companies belong to the ICT sector (30.8\%).
-However, many of them are represented by low percentages. For this reason, not having sufficiently
-representative samples, we have not considered particular data analysis and correlations by sectors.
+However, many of them are represented by low percentages. For this reason, missing sufficiently
+representative samples, we have not performed a particular data analysis and correlations by sectors.
 The 2020 pool is well balanced for the size of the organizations, in terms
@@
 belong to structures with less than 250 employees and, of these, 22.1\% have fewer than ten
-employees.
-The distribution of annual respondents' turnover also reflects the turnover of SMEs, as shown in Fig.~\ref{fig:7}.
-Regarding geographical distribution (Fig.~\ref{fig:8}), the responders' companies and organizations are located for the most part only on Italian territory (77.9\%), predominantly in northern Italy (43.3\%).
+employees. The distribution of annual respondents' turnover also reflects the turnover of SMEs,
+as shown in Fig.~\ref{fig:7}. Regarding geographical distribution (Fig.~\ref{fig:8}),
+the responders' companies and organizations are mostly located in the Italian territory (77.9\%),
+predominantly in northern Italy (43.3\%).
 
@@
 	\caption{Last annual available turnover of responders' companies and organizations~\cite{oad20}. 
-		 It should be noted that almost 20\% of the respondents cannot or do not want provide the turnover class to which they fall, stating that the confidentiality policies prevent to indicating a turnover class, even in an investigation completely anonymous.
+	  It should be noted that almost 20\% of the respondents cannot or do not want
+          to provide the turnover class to which they fall, asserting that the confidentiality
+          policies prevent them from indicating a turnover class,
+          even in an completely anonymous investigation.
 	}
@@
 Regarding the geographical distribution of attacks, Table.~\ref{tab:table4} shows
-that occasional and infrequent attacks are evenly distributed throughout the territory in 2019,
-with systematic attacks focused on northern Italy (where the companies with the highest turnover are present). 
-While, Table.~\ref{tab:table5} shows that already in the first months of 2020 there were several attacks.
+that occasional attacks were evenly distributed throughout the territory in 2019,
+with systematic attacks focused on northern Italy
+(where the companies with the highest turnover are present). 
+While Table.~\ref{tab:table5} shows that, already in the first months of 2020,
+there were several attacks.
 
@@
 
-	\caption{Distribution of digital attacks (detected by the respondents) wrt. to geographic area in 2019. The percentages of central Italy are 
-	very different from the others because the organizations responding to that geographical area were numerically lower than the rest 
+	\caption{Distribution of digital attacks (detected by the respondents) wrt.\ geographical area, in 2019. The percentages of central Italy are 
+	very different from the others because the organizations responding from that geographical area were much fewer than the rest 
 	of the survey participant (Fig.\ref{fig:8}).}									
@@
 
-	\caption{Distribution of digital attacks (detected by the respondents) wrt. to geographic area in 1Q-2020.}									
+	\caption{Distribution of digital attacks (detected by the respondents) wrt.\ geographical area, in 1Q-2020.}									
 	\label{tab:table5}									
@@
 (Table~\ref{tab:table2}) and for the cyber-terrorism (Table~\ref{tab:table3}). 
-Fraudulent transactions all Italyin 2019 were blocked for a total value of \euro 21.3 million and \euro 18 million were recovered; in the first quarter of 
+Fraudulent transactions were blocked in all Italy in 2019 for a total value
+of \euro 21.3 million and \euro 18 million were recovered; in the first quarter of 
 2020, fraudulent transactions for \euro 20.2 million were blocked,
